### 2009-03-30 eisa/inst/doc/ISA_tutorial.Rnw
Add the biology related files as a separate package.

\documentclass{article}
\usepackage{ragged2e}

\newcommand{\Rfunction}[1]{\texttt{#1}}
\newcommand{\Rpackage}[1]{\texttt{#1}}
\newcommand{\Rclass}[1]{\texttt{#1}}
\newcommand{\filename}[1]{\texttt{#1}}

\begin{document}

\title{The Iterative Signature Algorithm}
\author{G\'abor Cs\'ardi}
\maketitle

\RaggedRight

\section{Introduction}

The Iterative Signature Algorithm (ISA) is biclustering method. The
input of a biclustering method a matrix and its output is a set of
biclusters that fulfill some criteria; a bicluster is a block of the
potentially reordered input matrix.

Most commonly, this algorithm is used on microarray expression data,
to find gene sets that are coexpressed across a subset of the original
samples. In the original ISA paper the biclusters are called
transription modules (TM), we will often refer them under this name in
the following.

In this tutorial we will use the data in the \Rpackage{ALL}
package. This is a data set from a clinical trial in acute
lymphoblastic leukemia and it contains 128 samples altogether.

\section{Preparing the data}

\subsection{Loading the data}

First, we load the required packages and the data to analyze.
<<>>=
library(isa)
library(ALL)
library(hgu95av2.db)
data(ALL)
@

\subsection{Filtering the genes}

<<>>=
library(genefilter)
f1 <- pOverA(0.25, log2(100))
f2 <- function(x) (IQR(x) > 0.5)
ff <- filterfun(f1, f2)
selected <- genefilter(ALL, ff)
ALL2 <- ALL[selected,]
@

\subsection{Generating seeds for the ISA}

<<>>=
g <- generate.seeds(count=100, length=nrow(ALL2), gs=2/nrow(ALL2))
@

\subsection{Normalize the data}

<<>>=
ALL2.normed <- isa.normalize(ALL2)
@

\section{Running the ISA}

<<>>=
thr.gene <- 2.7
thr.cond <- 1.4
@

<<>>=
biclusts <- isa(ALL2.normed, thr.gene=thr.gene, thr.cond=thr.cond)
@

\section{Visualizing the results}

\section{Session information}

The version number of R and packages loaded for generating this
vignette were:

<<results=tex,echo=FALSE>>=
gsub(";", "; ", gsub("_", "\\\\_", sub("Locale:[ ]*\\\\verb\\|(.*)\\|", 
                                       "\\\\texttt\\{\\1\\}", 
                                       toLatex(sessionInfo()))))
@

\end{document}