### 2009-04-01 isa/inst/doc/ISA_tutorial.Rnw
Started writing the isa vignette.

% \VignetteIndexEntry{The Iterative Signature Algorithm}
\documentclass{article}
\usepackage{ragged2e}

\newcommand{\Rfunction}[1]{\texttt{#1}}
\newcommand{\Rpackage}[1]{\texttt{#1}}
\newcommand{\Rclass}[1]{\texttt{#1}}
\newcommand{\Rargument}[1]{\textsl{#1}}
\newcommand{\filename}[1]{\texttt{#1}}

\begin{document}

\setkeys{Gin}{width=\textwidth}

\title{The Iterative Signature Algorithm}
\author{G\'abor Cs\'ardi}
\maketitle

% \RaggedRight

\tableofcontents

\section{For the impatient}

To run the typical ISA work flow with default parameters on your data
matrix, just load the \Rpackage{isa} package and call the
\Rfunction{isa} function with your matrix as the single argument. The
return value of \Rfunction{isa} is a named list, the members
\texttt{rows} and \texttt{columns} contain the biclusters ISA have
found. Every bicluster is given by a pair of columns from
\texttt{rows} and \texttt{columns} (i.e. the first columns define the
first bicluster, etc.) and the elements the biclusters are the
non-zero elements in the columns of \texttt{rows} and
\texttt{columns}.

Please continue reading for a less dense tutorial.

\section{Introduction}

The Iterative Signature Algorithm (ISA) is biclustering method. Its 
input is a matrix and its output is a set of biclusters: blocks of the
potentially reordered input matrix, that fulfill some predefined
criteria. A biclustering algorithm tries to find
blocks that are different from the rest of the matrix, e.g. the values
covered by the bicluster are all above or below the background.

The ISA is developed to find biclusters (or modules as most of the ISA
papers call them) that have correlated rows and columns. More
precisely, the rows in the bicluster need to be only correlated across
the columns of the bicluster and vice versa.

Fig.~\ref{fig:twoclusters} shows possibly the simplest
example of a (rather artificial) data matrix with very strong modular
structure. It is a $20\times 20$ matrix and has two correlated blocks,
each of them of size $10\times 10$.
\begin{figure}
<<echo=FALSE,fig=TRUE,width=6,height=3>>=
library(isa)
library(Matrix)
data <- data2 <- isa.in.silico(20,20,2,10,10)[[1]]
perm1 <- sample(seq_len(nrow(data)))
data2 <- data2[perm1,]
perm2 <- sample(seq_len(ncol(data)))
data2 <- data2[,perm2]

alldata <- structure(c(data2, data), dim=c(nrow(data),ncol(data),2), 
                     dimnames=list(NULL,NULL,c("Original matrix",
                     "Reordered matrix")))

rx <- range(alldata, finite = TRUE)
nn <- 100
n0 <- min(nn, max(0, round((0 - rx[1])/(rx[2] - rx[1]) * nn)))
col.regions <- c(colorRampPalette(c("blue3", "gray80"))(n0), 
                 colorRampPalette(c("gray75", "red3"))(nn - n0))

lp <- levelplot(alldata, xlab="", ylab="", 
                scale=list(y=list(draw=FALSE),
                  x=list(draw=FALSE)),
                col.regions=col.regions, between=list(x=0.5))
print(lp)
@ 
\caption{An artificial data matrix, on the left. On the right the
  reordered data matrix with two blocks.}
\label{fig:twoclusters}
\end{figure}

\section{How ISA works}

Before showing an actual ISA tool chain, a few words about how the
algorithm works are in order.

\subsection{ISA iteration}

ISA works in an iterative way. For an $E (m\times n)$
input matrix it starts from seed vector $r_0$, which is
typically a sparse 0/1 vector of length $m$. This defines a set of
rows in $E$. Then $E'$ is multiplied by $r_0$ and the
result is thresholded.

The thresholding is an important step of the ISA, without thresholding
ISA would be equivalent to a (not too effective) numerical singular
value decomposition (SVD). Currently thresholding is done by
calculating the mean and standard deviation of the vector and keeping
only elements that are further than a given number of standard
deviations from the mean. Based on the \Rargument{direction} parameter,
this means 1) keeping values that are significantly higher than the
mean (``\emph{up}''), significantly lower (``\emph{down}'') or both
(``\emph{updown}'').

The thresholded vector $c_0$ is the (column)
\emph{signature} of $r_0$. Then the (row) signature of
$c_0$ is calculated, $E$ is multiplied by $c_0$ and then thresholded
to get $r_1$.

This iteration is performed until it converges, i.e. $r_i$
and $r_{i-1}$ are \emph{close}, and $c_i$ and
$c_{i-1}$ are also close. The convergence criteria,
i.e. what \emph{close} means is by default defined by high Pearson
correlation.

It is very possible that the ISA finds the same modules more than once;
two or more seeds might converge to the same module. The function
\Rfunction{isa.unique} eliminates every module from the result of 
\Rfunction{isa.iterate} that is very similar (in terms of
Pearson correlation) to the one that was already found before it.

It might be also apparent from the description of ISA, that the
biclusters are soft, i.e. they might have an overlap in their rows,
columns, or both. It is also possible that some rows and/or
columns of the input matrix are not found to be part of any ISA
biclusters. Depending on the stringency parameters in the
thresholding, it might even happen that ISA does not find any
biclusters.

\subsection{Parameters}

The two main parameters of ISA are the two thresholds (one for the
rows and one for the columns). They basically define the stringency of
the modules. If the row threshold is high, then the modules will have
very similar rows. If it is mild, then modules will be bigger, with
less similar rows than in the first case.

\subsection{Random seeding and smart seeding}

By default (i.e. if the \Rfunction{isa} function is used) the ISA is
performed from random sparse starting seeds, generated by the
\Rfunction{generate.seeds} function. This way the algorithm is 
completely unsupervised, but also stochastic: it might give different
results for different runs.

It is possible to use non-random seeds as well, if you have some
knowledge about the data or are interested in a particular subset of
rows/columns, then you can feed in your seeds into the
\Rfunction{isa.iterate} function directly. In this case the
algorithm is deterministic, for the same seed you will always get the
same results. This can be considered as a semi-supervised approach.

\subsection{Normalization}

On in silico data we observed that ISA has the best performance if the
input matrix is normalized (see \Rfunction{isa.normalize}). The
normalization produces two matrices: $E_r$ and
$E_c$. $E_r$ is calculated by transposing $E$ and
centering and scaling its rows (see \Rfunction{scale}). $E_c$ is
calculated by centering and scaling the rows of $E$. $E_r$ is
used to calculate the column signature of rows and $E_c$ is used
to calculate the signature of the columns.

It is possible to use another normalization, then the user is
requested to supply the normalized input data in a named list,
including the two matrices of appropriate
dimensions. \texttt{Er} will be used for calculating the
signature of the rows, \texttt{Ec} the signature of the
columns. If you want to use the same matrix in both steps, then supply
it twice, the first one transposed.

\subsection{Robustness}
As ISA is an unsupervised algorithm, it may very well find some
modules, even if you feed in noise as an input matrix. To avoid these
spurious modules we defined a robustness measure, a single number for
a modules that gives how well the rows and the columns are
correlated.

It recommended that the user uses \Rfunction{isa.filter.robust} to
run ISA on the scrambled input matrix with the same threshold
parameters and then drop every module, which has a robustness score
lower than the highest robustness score among modules found in the
scrambled data.

\section{A simple work flow}

\section{A detailed work flow}

\subsection{Preparing the data}

\subsection{Running the ISA}

\subsection{Robustness of biclusters, filtering the results}

\subsection{Visualize the results}

\section{Features of ISA}

\subsection{Resilience to noise}

\subsection{Finding overlapping biclusters}

\section{Session information}

The version number of R and packages loaded for generating this
vignette were:

<<results=tex,echo=FALSE>>=
gsub(";", "; ", gsub("_", "\\\\_", sub("Locale:[ ]*\\\\verb\\|(.*)\\|", 
                                       "\\\\texttt\\{\\1\\}", 
                                       toLatex(sessionInfo()))))
@

\end{document}