### 2009-06-10 eisa/inst/doc/ISA_tutorial.Rnw
eisa vignette extended a bit.

\documentclass{article}
\usepackage{ragged2e}

\newcommand{\Rfunction}[1]{\texttt{#1}}
\newcommand{\Rpackage}[1]{\texttt{#1}}
\newcommand{\Rclass}[1]{\texttt{#1}}
\newcommand{\filename}[1]{\texttt{#1}}

\begin{document}

\title{The Iterative Signature Algorithm on Gene Expression data}
\author{G\'abor Cs\'ardi}
\maketitle

\RaggedRight

\section{Introduction}

The Iterative Signature Algorithm (ISA) is biclustering method. The
input of a biclustering method a matrix and its output is a set of
biclusters that fulfill some criteria; a bicluster is a block of the
potentially reordered input matrix.

Most commonly, this algorithm is used on microarray expression data,
to find gene sets that are coexpressed across a subset of the original
samples. In the original ISA paper the biclusters are called
transription modules (TM), we will often refer them under this name in
the following.

\section{Preparing the data}

\subsection{Loading the data}

First, we load the required packages and the data to analyze. ISA is
implemented in the \Rpackage{eisa} and \Rpackage{isa2} packages, see
Section{sec:isapackages} for a more elaborated summary about the two
packages. It is enough to load the \Rpackage{eisa} package,
\Rpackage{isa2} and other required packages are loaded automatically.
\Rpackage{eisa} 
<<load the packages>>=
library(eisa)
@

In this tutorial we will use the data in the \Rpackage{ALL}
package. This is a data set from a clinical trial in acute
lymphoblastic leukemia and it contains 128 samples altogether.
<<load the data>>=
library(ALL)
library(hgu95av2.db)
data(ALL)
@

\section{Simple ISA runs}

The simplest way to run ISA is to choose the two threshold parameters
and then call the \Rfunction{isa} function on the
\Rclass{ExpressionSet} object:
<<simple ISA>>=
thr.gene <- 2.7
thr.cond <- 1.4
set.seed(1) # to get the same results, always
biclusts <- isa(ALL, thr.gene=thr.gene, thr.cond=thr.cond)
@

This first applies a non-specific filter to the data set filters the
data set and then runs ISA from
\Sexpr{formals(eisa:::isa.ExpressionSet)$no.seeds} % $
random seeds (the default). See Section~\ref{sec:detailed-isa} if the
default parameters are not appropriate for you and need more control.

\section{Inspect the result}

The \Rfunction{isa} function returns an \Rclass{ISAModules} object. By
typing in its name we can get a brief summary of the results:
<<type in name of ISAModules object>>=
biclusts
@

There are various other \Rclass{ISAModules} methods that help to
access the modules themselves and the ISA parameters that were used
for the run.

% TODO: length, dim, featureNames, sampleNames, pData, 
% annotation, organism, indexing

The \Rfunction{getNoFeatures} function returns a numeric vector, the
number of features (probesets in our case) in each module. Similarly,
\Rfunction{getNoSamples} returns a numeric vector, the number of
samples in each module. 
<<number of features and samples>>=
getNoFeatures(biclusts)
getNoSamples(biclusts)
@

\Rfunction{getFeatureNames} lists the probes (more precisely,
the feature names coming from the \Rclass{ExpressionSet} object) in
the modules. It returns a list, here we just print the first entry in
the list.
<<list genes in modules>>=
getFeatureNames(biclusts)[[1]]
@

The \Rfunction{getSampleNames} function does the same for the samples.
Again, the sample names are taken from the \Rclass{ExpressionSet}
object that was passed to \Rfunction{isa}:
<<list conditions in modules>>=
getSampleNames(biclusts)[[1]]
@

Remember, that the ISA biclustering is not binary, every feature (and
similarly, every sample) has a score between -1 and 1; the further the
score is from zero the stronger the association between the feature
(or sample) and the module. If two features both have scores with the
same sign, then they are correlated, if the sign of their scores are
opposite, then they are anti-correlated. You can query the scores of
the features with the \Rfunction{getFeatureScores} function, and
similarly, the \Rfunction{getSampleScores} function queries the sample
scores. You can supply the modules you want to query as an optional
argument:
<<query scores>>=
getFeatureScores(biclusts, 3)
getSampleScores(biclusts, 3)
@

You can also query the scores in a matrix form, that is probably
better if you need many or all of them at the same time. The
\Rfunction{getFeatureMatrix} and \RFunction{getSampleMatrix} functions
are defined for this:
<<query all scores>>=
dim(getFeatureMatrix(biclusts))
dim(getSampleMatrix(biclusts))
@

Objects from the \Rclass{ISAModules} class store various information
about the ISA run and the convergence of the seeds. Information
associated with the individual seeds can be queried with the
\Rfunction{seedData} function, it returns a data frame, with as many
rows as the number of seeds and various seed-level information,
e.g. the number of iterations required for the seed to converge. See
the manual page of \Rfunction{isa} for details.
<<seed data>>=
seedData(biclusts)
@

The \Rfunction{runData} function returns additional information about
the ISA run, seed the \Rfunction{isa} manual page for details.
<<run data>>=
runData(biclusts)
@

\section{Enrichment calculations}

The \Rpackage{eisa} package provides some functions to perform
enrichment tests for the genes in the various modules against various
databases. These tests are usually simplified and less sophisticated
versions than the ones in the \Rpackage{Category}, \Rpackage{GOstats}
or \Rpackages{topGO} packages, but they are much faster and this is
important if we need to perform them for many modules.

\subsection{Gene Ontology}

To perform enrichment analysis against the Gene Ontology database, all
you have to do is to supply your \Rclass{ISAModules} object to the
\Rfunction{isa.GO} function.

\subsection{KEGG Pathway Database}

\subsection{Chromosomes}

\subsection{Predicted $\mu$RNA targets 
  from the TargetScan database}

\subsection{Transcription regulation from the DBD database}

\section{Visualizing the results}

\subsection{Generating a HTML summary for the modules}

\section{The \Rpackage{isa2} and \Rpackage{eisa} packages}%
\label{sec:isapackages}

ISA and its companion functions for visualization, functional
enrichment calculation, etc. are distributed in two separate packages,
\Rpackage{isa2} and \Rpackage{eisa}.

\Rpackage{isa2} contains the implementation of ISA itself, and
\Rpackage{eisa} specifically deals with supplying expression data to
\Rpackage{isa2} and visualizing the results.

The two packages provide some functions that share the same name, and
the right one is selected based on the arguments supplied. E.g. both
\Rpackage{isa2} and \Rpackage{eisa} has a function called
\Rfunction{isa}. It the first argument of \Rfunction{isa} is an
\Rclass{ExpressionSet}, then the \Rpackage{eisa} version is called. If
the first argument is a matrix, then the \Rpackage{isa} version is
called.

This also means that there are two manual pages for the two
\Rfunction{isa} functions. I.e. if one types `\verb+?isa+', (and both
\Rpackage{isa2} and \Rpackage{eisa} are loaded, then one
has to choose between the two versions. To avoid this, the
\Rpackage{eisa} manual pages can also be referenced by supplying and
`\verb+e+' prefix: `\verb+?eisa+' or `\verb+?eisa.unique+'.

\section{Finer control over ISA parameters}%
\label{sec:detailed-isa}

\section{Session information}

The version number of R and packages loaded for generating this
vignette were:

<<results=tex,echo=FALSE>>=
gsub(";", "; ", gsub("_", "\\\\_", sub("Locale:[ ]*\\\\verb\\|(.*)\\|", 
                                       "\\\\texttt\\{\\1\\}", 
                                       toLatex(sessionInfo()))))
@

\end{document}