### 2009-10-30 eisa/inst/doc/EISA_module_trees.Rnw

% \VignetteIndexEntry{ISA module trees}
\documentclass{article}
\usepackage{ragged2e}
\usepackage{url}

\newenvironment{narrow}[2]{%
  \begin{list}{}{%
  \setlength{\topsep}{0pt}%
  \setlength{\leftmargin}{#1}%
  \setlength{\rightmargin}{#2}%
  \setlength{\listparindent}{\parindent}%
  \setlength{\itemindent}{\parindent}%
  \setlength{\parsep}{\parskip}}%
\item[]}{\end{list}}

\newcommand{\Rfunction}[1]{\texttt{#1()}}
\newcommand{\Rpackage}[1]{\texttt{#1}}
\newcommand{\Rclass}[1]{\texttt{#1}}
\newcommand{\Rargument}[1]{\textsl{#1}}
\newcommand{\filename}[1]{\texttt{#1}}
\newcommand{\variable}[1]{\texttt{#1}}

\SweaveOpts{cache=TRUE}
\SweaveOpts{prefix.string=plot}

\setlength{\parindent}{2em}

\begin{document}

\title{Transcription Module Trees from Gene Expression Data}
\author{G\'abor Cs\'ardi}
\maketitle

\tableofcontents

\RaggedRight

<<set width,echo=FALSE,print=FALSE>>=
options(width=60)
options(continue=" ")
@

\section{Introduction}

The Iterative Signature Algorithm (ISA) finds transcription modules in
gene expression data. A module is a block of the reordered gene
expression matrix.

The ISA has two threshold parameters: the gene (or feature) threshold
and the conditions (or sample) threshold. These essentially specify
the stringency of the biclustering. If the gene threshold is high,
then modules will have less genes, only the one that are consistently
far above or below the mean expression level will be included. The
same is true for the sample threshold and the samples in the modules.

It is possible to run the ISA using different threshold parameters and
define a hierarchical organization of the found modules. We call such
a description a sweep graph or sweep tree. The vertices (or nodes) of
the sweep graph are the modules and there is a directed edge from
module A to module B if the ISA converges to module B, when started
from module A, and the thresholds that were used to find module B are
used.

A sweep tree can be generated by keeping one threshold parameter
fixed, and vary the other one. We first find modules at a stringent
threshold value and then check where these converge at less stringent
thresholds.

This document shows how it is possible to create a sweep tree with the
\Rpackage{eisa} package and the tree can be visualized.

\section{Running the ISA}

We will use the acute lymphoblastic leukemia data set from the
\Rpackage{ALL} package.

<<load data,cache=FALSE>>=
library(ALL)
library(hgu95av2.db)
library(genefilter)
library(eisa)
library(GO.db)
data(ALL)
@

<<filter limits>>=
varLimit <- 0.5
kLimit <- 4
ALimit <- 5
@

First we filter the genes and keep only the ones that have an
expression level of at least \Sexpr{ALimit} in at least \Sexpr{kLimit}
samples; and that have a variance of at least \Sexpr{varLimit} across
the samples.

<<filtering the data>>=
flist <- filterfun(function(x) var(x)>varLimit, kOverA(kLimit,ALimit))
ALL.filt <- ALL[genefilter(ALL, flist), ]
@

We remove the T-cell samples as well, and use only the B-cell leukemia
samples.

<<keep B cell data only>>=
ALL.filt2 <- ALL.filt[, grepl("^B", ALL.filt$BT)]
@

Next, we run the ISA, with keeping the condition threshold fixed, but
varying the gene threshold.

<<run ISA>>=
set.seed(2)
modules <- ISA(ALL.filt2, flist=NA, thr.gene=seq(2,4,by=0.5), thr.cond=1)
@

\section{Generating the module tree}

The \Rfunction{ISASweep} function can be used to create a sweep tree
from a set of ISA modules. This essentially means finding the edges
between the modules and occasionally involves finding new modules as
well.

<<sweep>>=
modules2 <- ISASweep(ALL.filt2, modules)
@

\Rfunction{ISASweep} returns an \Rclass{ISAModules} object, but adds
some new seed data that defines the edges of the sweep graph, plus
possibly some new modules:
<<sweep result>>=
modules
modules2
colnames(seedData(modules))
colnames(seedData(modules2))
@

The \texttt{father} column contains pointers to the parent vertices in
the graph and the \texttt{level} column gives levels defined based on
the gene thresholds, higher levels indicate less stringent thresholds.

This data can be converted into a graph and then plotted using the
\Rpackage{igraph} package. The \Rfunction{ISASweepGraph} function
creates the graph.

<<getting the graph>>=
G <- ISASweepGraph(modules2)
@

$G$ has a lot of attributes:
<<G>>=
list.graph.attributes(G)
list.vertex.attributes(G)
@

The \texttt{layout} graph attribute contains a two-column matrix with
carefully calculated coordinates for plotting the graph. The
\texttt{width} and \texttt{height} graph attributes contain the
suggested size of the plot in inches.
<<G size>>=
G$width
G$height
@

The \texttt{id} vertex attribute contains the ids of the modules. 
The other vertex attributes essentially define the look of the graph,
when it is plotted.

See the documentation of the \Rpackage{igraph} package for more about
igraph graph objects.

\section{Plotting the sweep tree}

\subsection{Simple plots}

The \Rfunction{ISASweepGraphPlot} function can be used to plot the
graph returned by \Rfunction{ISASweepGraph}. We will use the suggested
size for the plot, and leave all plotting options at their default
values. The result is in Fig.~\ref{fig:moduletree}.

<<plotting,eval=FALSE>>=
X11(width=G$width, height=G$height)
@ 
<<plotting2,eval=FALSE>>=
ISASweepGraphPlot(G)
@

\begin{figure}
\begin{narrow}{-1cm}{-1cm}
\setkeys{Gin}{width=\linewidth}
<<plot,fig=TRUE,width=10,height=10,echo=FALSE>>=
ps.options(fonts=c("serif", "mono"))
<<plotting2>>
@ 
\caption{An ISA module tree. Each rectangle symbolizes a transcription
  module, modules in the same column were found at the same gene
  threshold. An arrow between two modules indicates that one is the
  ``generalization'' of the other, in the sense that ISA converges to
  the more general (=bigger) module at the milder threshold.}
\label{fig:moduletree}
\end{narrow}
\end{figure}

\subsection{Customized plots}

Let us show a little example on how to customize the module tree
plot. We will add some annotation to the module, based on Gene
Ontology enrichment calculations. Let us perform the hypergeometric
test for the enrichment calculation. The $p$-values are automatically
corrected using the Benjamini-Hochberg method.

<<enrichment>>=
GO <- ISAGO(modules2)
@

We will color the vertices according to their enrichment in the
Biological Process GO ontology. We also create some labels for them,
the abbreviated GO terms will be used for this.

<<customize>>=
p.bp <- sapply(summary(GO$BP), function(x) as.integer(-log10(x$Pvalue[1])))
c.bp <- d.bp <- sapply(sigCategories(GO$BP), function(x) x[1])
d.bp[!is.na(c.bp)] <- sapply(mget(na.omit(c.bp), GOTERM), Term)
d.bp <- abbreviate(d.bp, 6)
colbar <- hcl(h=260, c=35, l=seq(30, 100, length=20))
colbar <- c("#FFFFFF", rev(colbar))
col <- colbar[p.bp]
@

<<plotting3,eval=FALSE>>=
X11(width=G$width, height=G$height)
@

We are ready to plot the sweep tree now. We put the abbreviated GO
terms to the top left corner, and the size of the module (i.e. number
of genes and number of samples) in the top right corner. We also
increase the height of the vertices a bit.

<<plotting4,eval=FALSE>>=
ISASweepGraphPlot(G, vertex.color=col, vertex.size2=50, 
                  vertex.label.topleft=d.bp,
                  vertex.label.topright=paste(V(G)$noFeatures, sep=",", 
                    V(G)$noSamples))
@

Finally, we add a key for the abbreviated names at the top right
corner. The result is in Fig.~\ref{fig:moduletree2}.

<<key,eval=FALSE>>=
key <- na.omit(d.bp)[unique(names(na.omit(d.bp)))]
key2 <- paste(key, sep=": ", names(key))
legend("topright", key2)
@

\begin{figure}
\begin{narrow}{-1cm}{-1cm}
\setkeys{Gin}{width=\linewidth}
<<plot2,fig=TRUE,width=10,height=10,echo=FALSE>>=
<<plotting4>>
<<key>>
@ 
\caption{The same module tree, but with some annotation added. The
  modules are colored according to their Gene Ontology Biological
  Process enrichment $p$-values, darker colors correspond to more
  significant enrichment. The enriched GO terms and the sizes of the
  modules were also added.}
\label{fig:moduletree2}
\end{narrow}
\end{figure}

\section{More information}

For more information about the ISA, please see the references
below. The ISA homepage at
\url{http://www.unil.ch/cbg/homepage/software.html} has example data
sets, and all ISA related tutorials and papers.

\section{Session information}

The version number of R and packages loaded for generating this
vignette were:

<<sessioninfo,results=tex,echo=FALSE>>=
toLatex(sessionInfo())
@

\bibliographystyle{apalike}
\bibliography{EISA}
\nocite{*}

\end{document}