### 2009-10-30 eisa/inst/doc/EISA_tutorial.Rnw

% \VignetteIndexEntry{The Iterative Signature Algorithm for Gene Expression Data}
\documentclass{article}
\usepackage{ragged2e}
\usepackage{url}

\newcommand{\Rfunction}[1]{\texttt{#1()}}
\newcommand{\Rpackage}[1]{\texttt{#1}}
\newcommand{\Rclass}[1]{\texttt{#1}}
\newcommand{\Rargument}[1]{\textsl{#1}}
\newcommand{\filename}[1]{\texttt{#1}}
\newcommand{\variable}[1]{\texttt{#1}}

\SweaveOpts{cache=TRUE}
\SweaveOpts{prefix.string=plot}

\setlength{\parindent}{2em}

\begin{document}

\title{The Iterative Signature Algorithm for Gene Expression data}
\author{G\'abor Cs\'ardi}
\maketitle

\tableofcontents

\RaggedRight

<<set width,echo=FALSE,print=FALSE>>=
options(width=60)
options(continue=" ")
@

\section{Introduction}

The Iterative Signature Algorithm (ISA) is biclustering method. The
input of a biclustering method a matrix and the output is a set of
biclusters that fulfill some criteria. A bicluster is a block of the
potentially reordered input matrix.

Most commonly, this algorithm is used on microarray expression data,
to find gene sets that are coexpressed across a subset of the original
samples. In the ISA papers the biclusters are called
transription modules (TM), we will often refer them under this name in
the following.

This tutorial specifically deals with the modular analysis of gene
expression data. Section~\ref{sec:isa} gives a short summary of how
ISA works. If you need more information of the underlying math
or want to apply it to other data, then please see the referenced
papers, the vignette titled ``The Iterative Signature Algorithm'' in
the \Rpackage{isa2} R package, or the ISA homepage at
\url{http://www.unil.ch/cbg/homepage/software.html}.

\section{Preparing the data}

\subsection{Loading the data}

First, we load the required packages and the data to analyze. ISA is
implemented in the \Rpackage{eisa} and \Rpackage{isa2} packages, see
Section~\ref{sec:isapackages} for a more elaborated summary about the
two packages. It is enough to load the \Rpackage{eisa} package,
\Rpackage{isa2} and other required packages are loaded automatically:
<<load the packages,cache=FALSE>>=
library(eisa)
@

In this tutorial we will use the data in the \Rpackage{ALL}
package. 
<<load the data,cache=FALSE>>=
library(ALL)
library(hgu95av2.db)
library(affy)
data(ALL)
@
This is a data set from a clinical trial in acute
lymphoblastic leukemia and it contains \Sexpr{ncol(ALL)} samples
altogether.

\section{Simple ISA runs}

The simplest way to run ISA is to choose the two threshold parameters
and then call the \Rfunction{isa} function on the
\Rclass{ExpressionSet} object. The threshold parameters tune the size
of the modules, less stringent (i.e. smaller) values result bigger,
less correlated modules. The optimal values depend on your data and
some experimentation is needed to determine them.

Since running ISA might take a couple of minutes and the results 
depend on the random number generator used, the ISA run is commented out 
and we just load a precomputed set of modules in this vignette.
<<simple ISA>>=
thr.gene <- 2.7
thr.cond <- 1.4
set.seed(1) # to get the same results, always
# modules <- ISA(ALL, thr.gene=thr.gene, thr.cond=thr.cond)
data(ALLModulesSmall)
modules <- ALLModulesSmall
@

This first applies a non-specific filter to the data set and then runs
ISA from \Sexpr{formals(eisa:::ISA)$no.seeds} % $
random seeds (the default). See Section~\ref{sec:detailed-isa} if the
default parameters are not appropriate for you and need more control.

\section{Inspect the result}

The \Rfunction{isa} function returns an \Rclass{ISAModules} object. By
typing in its name we can get a brief summary of the results:
<<type in name of ISAModules object>>=
modules
@

There are various other \Rclass{ISAModules} methods that help to
access the modules themselves and the ISA parameters that were used
for the run.

Calling \Rfunction{length} on \variable{modules} returns the number of
ISA modules in the set, \Rfunction{dim} gives the dimension of the
input expression matrix: the number of features (after the filtering)
and the number of samples:
<<accessors 1>>=
length(modules)
dim(modules)
@

Functions \Rfunction{featureNames} and \Rfunction{sampleNames} return
the names of the features and samples, just like the functions with
the same name for an \Rclass{ExpressionSet}:
<<accessors 2>>=
featureNames(modules)[1:5]
sampleNames(modules)[1:5]
@

The \Rfunction{getNoFeatures} function returns a numeric vector, the
number of features (probesets in our case) in each module. Similarly,
\Rfunction{getNoSamples} returns a numeric vector, the number of
samples in each module. \Rfunction{pData} returns the
phenotype data of the expression set as a data frame. The
\Rfunction{getOrganism} function returns the scientific name of the
organism, \Rfunction{annotation} the name of the chip. For the former
the appropriate annotation package must be installed.
<<number of features and samples>>=
getNoFeatures(modules)
getNoSamples(modules)
colnames(pData(modules))
getOrganism(modules)
annotation(modules)
@

The double bracket indexing operator (`\verb+[[+') can be used to
select some modules from the complete set, the result is another,
smaller \Rclass{ISAModules} object. The following selects the first
five modules.
<<indexing>>=
modules[[1:5]]
@

The single bracket indexing operator can be used to restrict an
\Rclass{ISAModules} object to a subset of features and/or
samples. E.g. selecting all features that map to a gene on
chromosome~1 can be done with 
<<indexing 2>>=
chr <- get(paste(annotation(modules), sep="", "CHR"))
entrez <- sapply(mget(featureNames(modules), chr), 
                 function(x) "1" %in% x)
modules[entrez,]
@

Similarly, selecting all B-cell samples can be performed with
<<indexing 3>>=
modules[ ,grep("^B", pData(modules)$BT)]
@

\Rfunction{getFeatureNames} lists the probesets (more precisely,
the feature names coming from the \Rclass{ExpressionSet} object) in
the modules. It returns a list, here we just print the first entry in
the list.
<<list genes in modules>>=
getFeatureNames(modules)[[1]]
@

The \Rfunction{getSampleNames} function does the same for the samples.
Again, the sample names are taken from the \Rclass{ExpressionSet}
object that was passed to \Rfunction{isa}:
<<list conditions in modules>>=
getSampleNames(modules)[[1]]
@

ISA biclustering is not binary, every feature (and
similarly, every sample) has a score between -1 and 1; the further the
score is from zero the stronger the association between the feature
(or sample) and the module. If two features both have scores with the
same sign, then they are correlated, if the sign of their scores are
opposite, then they are anti-correlated. You can query the scores of
the features with the \Rfunction{getFeatureScores} function, and
similarly, the \Rfunction{getSampleScores} function queries the sample
scores. You can supply the modules you want to query as an optional
argument:
<<query scores>>=
getFeatureScores(modules, 3)
getSampleScores(modules, 3)
@

You can also query the scores in a matrix form, that is probably
better if you need many or all of them at the same time. The
\Rfunction{getFeatureMatrix} and \Rfunction{getSampleMatrix} functions
are defined for this:
<<query all scores>>=
dim(getFeatureMatrix(modules))
dim(getSampleMatrix(modules))
@

Objects from the \Rclass{ISAModules} class store various information
about the ISA run and the convergence of the seeds. Information
associated with the individual seeds can be queried with the
\Rfunction{seedData} function, it returns a data frame, with as many
rows as the number of seeds and various seed-level information,
e.g. the number of iterations required for the seed to converge. See
the manual page of \Rfunction{ISA} for details.
<<seed data>>=
seedData(modules)
@

The \Rfunction{runData} function returns additional information about
the ISA run, seed the \Rfunction{isa} manual page for details.
<<run data>>=
runData(modules)
@

\section{Enrichment calculations}

The \Rpackage{eisa} package provides some functions to perform
enrichment tests for the gene sets corresponding to the ISA modules
against various databases. These tests are usually simplified and less
sophisticated versions than the ones in the \Rpackage{Category},
\Rpackage{GOstats} or \Rpackage{topGO} packages, but they are much
faster and this is important if we need to perform them for many
modules.

\subsection{Gene Ontology}

To perform enrichment analysis against the Gene Ontology database, all
you have to do is to supply your \Rclass{ISAModules} object to the
\Rfunction{ISAGO} function.
<<GO enrichment>>=
GO <- ISAGO(modules)
@

The \Rfunction{ISAGO} function requires the annotation package on the
chip, e.g. for the ALL data, the \Rpackage{hgu95av2.db} package is
required.

The \variable{GO} object is a list with three elements, these
correspond to the GO ontologies, they are: biological function,
cellular component and molecular function, in this order.
<<list GO result>>=
GO
@

We can see the number of categories tested, this is different for each
ontology, as they have different number of terms. The gene universe
size is also different, because it contains only genes that have at
least one annotation in the given category.

For extracting the results themselves, the \Rfunction{summary}
function can be used, this converts them to a simple data frame. A
$p$-value limit can be supplied to \Rfunction{summary}. Note, that
since \Rfunction{ISAGO} calculates enrichment for many gene sets
(i.e. for all biclusters), \Rfunction{summary} returns a list of data
frames, one for each bicluster. A table for the first module:
<<GO summary>>=
summary(GO$BP, p=0.001)[[1]][,-6]
@

We omitted the sixth column of the result, because it is very wide and
would look bad in this vignette. This column is called
\variable{drive} and lists the Entrez IDs of the genes that are in the
intersection of the bicluster and the GO category; or in other
words, the genes that drive the enrichment.
These genes can also be obtained with the
\Rfunction{geneIdsByCategory} function. The following returns the
genes in the first module and the third GO BP category. (The GO
categories are ordered according to the enrichment $p$-values, just
like in the output of \Rfunction{summary}.)
<<GO gene ids by category>>=
geneIdsByCategory(GO$BP)[[1]][[3]]
@

You can use the \Rpackage{GO.db} package to obtain more information
about the enriched GO categories.
<<GO info,cache=FALSE>>=
sigCategories(GO$BP)[[1]]
library(GO.db)
mget(sigCategories(GO$BP)[[1]], GOTERM)
@

In addition, the following functions are implemented to work on the
objects returned by \Rfunction{ISAGO}: \Rfunction{htmlReport},
\Rfunction{pvalues}, \Rfunction{geneCounts}, \Rfunction{oddsRatios},
\Rfunction{exprectedCounts}, \Rfunction{universeCounts},
\Rfunction{universeMappedCount}, \Rfunction{geneMappedCount},
\Rfunction{geneIdUniverse}. These functions do essentially the same as
they counterparts for \Rclass{GOHyperGResult} objects, see the
documentation of the \Rpackage{GOstats} package. The only difference
is, that since here we are testing a list of gene sets (=biclusters),
they calculate the results for all gene sets and usually return lists.

\subsubsection{Multiple testing correction}

By default, the \Rfunction{ISAGO} function performs multiple testing
correction using the Holm method, this can be changed via the
\texttt{correction} and \texttt{correction.method} arguments. See the
manual page of the \Rfunction{ISAGO} function for details, and also
the \Rfunction{p.adjust} function for the possible multiple testing
correction schemes.

\subsection{KEGG Pathway Database}

Enrichment calculation against the KEGG pathway database goes
essentially the same way as for the Gene Ontology, this time we use
the \Rfunction{ISAKEGG} function:
<<KEGG enrichment,cache=FALSE>>=
KEGG <- ISAKEGG(modules)
KEGG
summary(KEGG, p=1e-3)[[1]]
library(KEGG.db)
mget(sigCategories(KEGG)[[1]], KEGGPATHID2NAME)
@

The functions mentioned in the Gene ontology enrichment section can be
used for KEGG, as well.

\subsection{Chromosomes}

The \Rpackage{eisa} includes a simple way to check whether the genes
in a bicluster are associated with a chromosome. See the
\Rfunction{ISACHR} function:
<<CHR enrichment>>=
CHR <- ISACHR(modules)
summary(CHR,p=0.05)[[2]][,-6]
@

The second bicluster has \Sexpr{summary(CHR,p=0.05)[[2]]$Count[1]} % $
genes on chromosome
\Sexpr{rownames(summary(CHR,p=0.05)[[2]])[1]}. Here is a list of the
genes:
<<CHR enrichment list>>=
unlist(mget(geneIdsByCategory(CHR)[[2]][[1]], org.Hs.egSYMBOL))
@

The functions mentioned in the Gene ontology enrichment section can be
used for chromosome enrichment, as well.

\subsection{Predicted $\mu$RNA targets 
  from the TargetScan database}

$\mu$RNAs are short RNA molecules that regulate gene expression.
TargetScan is a database of predicted target genes of $\mu$RNAs, for
several organisms. There are two R packages that incorporate this
data, one for human and another one for mouse, right now they can
be downloaded from
\url{http://www.unil.ch/cbg/homepage/software.html}.
The \Rpackage{targetscan.Hs.eg.db} package is for human, the
\Rpackage{targetscan.Mm.eg.db} package is for mouse.

The enrichment calculation itself is basically the same as for GO and
KEGG, but the \Rfunction{ISAmiRNA} function should be used:
<<miRNA enrichment>>=
if (require(targetscan.Hs.eg.db)) {
  miRNA <- ISAmiRNA(modules)
  summary(miRNA, p=0.1)[[7]]
}
@

%\subsection{Transcription regulation from the DBD database}

%TODO

\section{Visualizing the results}

Visualizing overlapping biclusters is a challenging task. We
show simple methods that usually visualize a single bicluster 
at a time. For some of these we will use the \Rpackage{biclust} R
package.

\subsection{The \Rpackage{biclust} package}%
\label{sec:biclust}

The \Rpackage{biclust} R package implements several biclustering
algorithms in a unified framework. It uses the class \Rclass{Biclust}
to store a set of biclusters. The standard R \Rfunction{as} function
can be used to convert ISA modules to a \Rclass{Biclust} object. This
requires the binarization of the modules, i.e. the ISA scores are
lost, they are converted to zeros and ones:
<<biclust,cache=FALSE>>=
library(biclust)
Bc <- as(modules, "Biclust")
Bc
@

\subsection{Image plots}

The easiest way to create a heatmap of a single module is to call the
\Rfunction{ISA2heatmap} function. You need to specify which module you
want to plot and also the \Rclass{ExpressionSet} object that is being
analyzed. Note that by default ISA normalizes the expression data
before running the module detection, but \Rfunction{ISA2heatmap} plots
the raw values by default, optionally scaled. 
If you want to plot the normalized values, then you need supply the
\Rargument{norm} argument to \Rfunction{ISA2heatmap}, see the manual
for details. The result is in Fig.~\ref{fig:heatmap}.
<<heatmap,results=hide,include=FALSE,eval=FALSE>>=
ISA2heatmap(modules, 1, ALL, norm="sample", scale="none")
@

\begin{figure}
\centering
<<heatmap-real,fig=TRUE,echo=FALSE,results=hide>>=
<<heatmap>>
@
\caption{Heatmap of the first ISA module, plotted via the
\Rfunction{ISA2heatmap} function. The expression values are
sample-wised normalized.}
\label{fig:heatmap}
\end{figure}

\Rfunction{ISA2heatmap} simply calls the \Rfunction{heatmap} function,
and passes additional arguments to it. See the manual of
\Rfunction{heatmap} for details.

You can also use the \Rpackage{biclust} package to create image
plots, by calling the \Rfunction{drawHeatmap} function. The result is
in Fig.~\ref{fig:drawHeatmap}.
<<heatmap biclust pre,echo=FALSE,results=hide>>=
if (require(Cairo)) {
  CairoPNG("plot-heatmap.png", 400, 400 )  
} else {
  png("plot-heatmap.png", 400, 400 )
}
@ 
<<heatmap>>=
drawHeatmap(exprs(ALL[featureNames(modules),]), Bc, number=1) 
@
<<heatmap post,echo=FALSE,results=hide>>=
dev.off() 
@ 
\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{plot-heatmap.png}
  \caption{Heatmap of the first transcription module, plotted with the
  \Rfunction{drawHeatmap} function from the \Rpackage{biclust}
  package.}
  \label{fig:drawHeatmap}
\end{figure}

\subsection{Profile plots}

Profile plots visualize the difference between the genes (or samples)
in the modules and the rest of the expression data. A profile plot
contains a line plot for every single gene (or sample) and the genes
that belong to the module have a different color, see
Fig.~\ref{fig:parallel}.
<<parallel.pre,echo=FALSE,results=hide>>=
if (require(Cairo)) {
  CairoPNG("plot-parallel.png", 400, 400 )
} else {
  png("plot-parallel.png", 400, 400 )
}
@ 
<<parallel>>=
profilePlot(modules, 2, ALL, plot="both")
@
<<parallel.post,echo=FALSE,results=hide>>=
dev.off()
@ 
\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{plot-parallel.png}
\caption{Profile plots for the second bicluster found by ISA.}
\label{fig:parallel}
\end{figure}

The \Rfunction{profile.plot} function has several options to set the
plot colors and styles, please see the manual for the details. This
function was inspired by the \Rfunction{parallelCoordinates} function
in the \Rpackage{biclust} package.

\subsection{Gene ontology tree plots}

The GO database is organized in a hierarchical fashion, in a tree-like
structure, where the broadest category sits in the root of the tree
and broader categories are subdivided into more specific
subcategories. But the GO is not exactly a tree, as the same category
can be the subcategory of more than one broader categories: e.g. the
``Golgi vesicle transport'' category is part of both
``vesicle-mediated transport'' and ``intracellular
transport''.

The \Rpackage{eisa} package provides functions to plot parts of
the GO graph that is related to a transcription module. The
\Rfunction{gograph} function creates an object that is a
representation of such a plot. Its input is a table with the GO
categories to plot and their enrichment $p$-values. (Additional
columns are silently ignored.) Here is how to use
it on the previously calculated enrichment scores:
<<goplot1>>=
goplot.2 <- gograph(summary(GO$BP, p=0.05)[[1]])
@ 
% $

\Rfunction{goplot} uses the \Rpackage{igraph} package to create a
graph with the associated meta data:
<<gograph object>>=
summary(goplot.2)
@ 
The \texttt{width} and \texttt{height} graph attributes contain the
suggested width and height of the graph in pixels, if plotted to a bitmap
device. (The graph attributes of an \Rpackage{igraph} graph can be
queried with the `\texttt{\$}' selector:
<<gograph object 2>>=
goplot.2$width
goplot.2$height
@

Let's plot the graph, we can do this with the \Rfunction{gographPlot}
function, see Fig.~\ref{fig:gotree}.
<<goplot2,eval=FALSE>>=
x11(width=10, height=10*goplot.2$height/goplot.2$width)
gographPlot(goplot.2)
@
\setkeys{Gin}{width=\textwidth}
<<goplot2-real,results=hide,include=FALSE,eval=FALSE,echo=FALSE>>=
ps.options(fonts=c("serif", "mono"))
gographPlot(goplot.2)
@

\begin{figure}
\centering
<<goplot2-real-real,width=10,height=5,fig=TRUE,results=hide,echo=FALSE,cache=FALSE>>=
<<goplot2-real>>
@
\caption{Part of the Gene ontology database, ``Biological Function''
ontology, that contains all GO terms enriched by a transcription module.}
\label{fig:gotree}
\end{figure}

Because the GO is not a tree, \Rfunction{gograph} ``unfolds''
it into a tree by including categories more than once, if needed. It
also abbreviates the names of the GO categories to make them fit on
the plot. The graph object contains the full name of the category as
well. The full and abbreviated names can be listed by querying the
appropriate vertex attributes of the graph. Here they are for the
first five categories:
<<goplot names>>=
V(goplot.2)$abbrv[1:5]
lapply(V(goplot.2)$desc[1:5], strwrap)
@ 
\Rfunction{gographPlot} colors the categories according to the
supplied enrichment $p$-values, the minus $\log_{10}$ $p$-value is
also added to the plot, see the bold blue numbers.

\subsection{Sample score plots}

In many studies, especially the case-control ones, it is useful to
plot the sample scores of a module. For example if the sample scores
significantly differ for two groups of samples (e.g. cases versus
controls), then the genes in the module can be used as discriminators
between the two groups.

The \Rfunction{condPlot} function can plot the sample scores,
potentially including the scores before the ISA filtering. Let us plot
the scores for the second module, the color code denotes B-cell
vs. T-cell leukemia (Fig.~\ref{fig:condplot}).

<<condplot,results=hide,include=FALSE,eval=FALSE>>=
col <- ifelse( grepl("^B", pData(modules)$BT), "darkolivegreen", "orange")
condPlot(modules, 2, ALL, col=col)
@
% $

\begin{figure}
\centering
<<condplot-real,fig=TRUE,width=10,height=4,echo=FALSE,results=hide>>=
<<condplot>>
@
\caption{Condition plot for the ALL data and the second module. Every
  sample is represented as a bar, and its ISA score is plotted. The
  scores are also printed at the top or bottom of the bars. The scores
  corresponding to the ISA thresholds are represented as horizontal
  lines. The dashed line is the mean of the condition scores. B-cell
  samples are dark green, T-cell samples are orange.
}
\label{fig:condplot}
\end{figure}

It is clear that the \Sexpr{getNoFeatures(modules)[2]} genes included
in this transcription module can separate B-cell and T-cell leukemia
samples.

\subsection{Generating a HTML summary for the modules}

The \Rfunction{ISAHTMLTable} function creates a summary of a set of
ISA modules, as an HTML table. The summary optionally includes the
results of the enrichment calculations as well. Fig.~\ref{fig:HTML}
shows a screenshot of the table created for the
\Sexpr{length(modules)} modules we have found earlier.
<<HTML1>>=
htmldir <- tempdir()
ISAHTMLTable(modules=modules, target.dir=htmldir, 
             GO=GO, KEGG=KEGG, CHR=CHR)
@ 
<<HTML2,eval=FALSE>>=
if (interactive()) {
  browseURL(URLencode(paste("file://", htmldir, "/maintable.html", sep="")))
}
@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{plot-HTMLTable.png}
\caption{A HTML summary table showing information for the first couple
of modules found in the ALL data set. The ``Thr.'' column shows the
ISA thresholds that were used to find the module; the next two columns
are the number of features (genes) and the number of samples
(conditions) in the modules. The other columns show the most
significantly enriched GO categories, KEGG pathways and
chromosomes. The small grey numbers are the enrichment $p$-value, the
number of genes expected to be hit by chance, the number of hits and
the size of the category (within the respective gene universe), in
this order.}
\label{fig:HTML}
\end{figure}

The Rfunction{ISAHTMLModules} generates a HTML page for every single
module with the following information on it: 
\begin{itemize}
\item An expression plot of the genes and samples in the module,
  including the ISA scores. This is done by calling \Rfunction{expPlot}.
\item Overlap plot of all modules, see \Rfunction{overlapPlot}.
\item  Gene Ontology tree plots for the enriched GO terms,
  separately for the three ontologies., These are produced by
  calling \Rfunction{gograph}.
\item Tables for the enriched Gene Ontology terms, separately for
  the three ontologies.
\item A table for the enriched KEGG pathways.
\item A table for the enriched miRNA families.
\item The list of genes in the module.
\item The list of samples in the module.
\item A condition plot, see the \Rfunction{condPlot} function.
\end{itemize}
<<HTML modules>>=
ISAHTMLModules(eset=ALL, modules=modules, target.dir=htmldir,
               GO=GO, KEGG=KEGG, CHR=CHR)
@ 
The rows of the summary table generated by \Rfunction{ISAHTMLTable}
link to the module pages generated by \Rfunction{ISAHTMLModules}.

Both the summary and the module pages can be generated with the
\Rfunction{ISAHTML} function.

\subsection{The ExpressionView package}

The previously mentioned visualization techniques focus on a single
module only. The \Rpackage{ExpressionView} package visualizes a set of
possibly overlapping biclusters together.

The \Rpackage{ExpressionView} package~\cite{ev} tries to reorder the
rows and columns of the expression matrix in such a way that the
genes/samples that appear together in a module, are placed right
beside/above each other. The package directly supports the
\Rclass{ISAModules} and \Rclass{ExpressionSet} classes, see its
documentation for details.

\section{How ISA works}%
\label{sec:isa}

\subsection{ISA iteration}

ISA works in an iterative way. For an $E (m\times n)$
input matrix it starts from a seed vector $r_0$, which is
typically a sparse 0/1 vector of length $m$. The non-zero elements in
the seed vector define a set of genes in $E$. Then the transposed of
$E$, $E'$ is multiplied by $r_0$ and the result is thresholded.

The thresholding is an important step of the ISA, without thresholding
ISA would be equivalent to a (not too effective) numerical singular
value decomposition (SVD) algorithm. Currently thresholding is done by
calculating the mean and standard deviation of the vector and keeping
only elements that are further than a given number of standard
deviations from the mean. Based on the \Rargument{direction} parameter,
this means keeping values that are significantly higher than the
mean (``\emph{up}''), or keeping the ones that are significantly lower
than the mean (``\emph{down}''); or keeping both
(``\emph{updown}'').

The thresholded vector $c_0$ is the (sample)
\emph{signature} of $r_0$. Then the (gene) signature of
$c_0$ is calculated, $E$ is multiplied by $c_0$ and then thresholded
to get $r_1$.

This iteration is performed until it converges, i.e. $r_{i-1}$
and $r_i$ are \emph{close}, and $c_{i-1}$ and
$c_i$ are also close. The convergence criteria,
i.e. what \emph{close} means, is by default defined by high Pearson
correlation.

It is very possible that the ISA finds the same module more than once;
two or more seeds might converge to the same module. The function
\Rfunction{ISAUnique} eliminates every module from the result of 
\Rfunction{ISAIterate} that is very similar (in terms of
Pearson correlation) to the one that was already found before.

It might be also apparent from the description of ISA, that the
biclusters are soft, i.e. they might have an overlap in their genes,
samples, or both. It is also possible that some genes and/or
samples of the input matrix are not found to be part of any ISA
biclusters. Depending on the stringency parameters in the
thresholding (i.e. how far the values should be from the mean), it
might even happen that ISA does not find any biclusters.

\subsection{Parameters}

The two main parameters of ISA are the two thresholds (one for the
genes and one for the samples). They basically define the stringency of
the modules. If the gene threshold is high, then the modules will have
very similar genes. If it is mild, then modules will be bigger, with
less similar genes than in the first case. The same applies to the
sample threshold and the samples of the modules.

\subsection{Random seeding and smart seeding}

By default (i.e. if the \Rfunction{ISA} function is used) the ISA is
performed from random sparse starting seeds, generated by the
\Rfunction{generate.seeds} function. This way the algorithm is 
completely unsupervised, but also stochastic: it might give different
results for different runs.

It is possible to use non-random seeds as well. If you have some
knowledge about the data or are interested in a particular subset of
genes/samples, then you can feed in your seeds into the
\Rfunction{ISAIterate} function directly. In this case the
algorithm is deterministic, for the same seed you will always get the
same results. Using smart (i.e. non-random) seeds can be considered as
a semi-supervised approach. We show an example of using smart seeds in
Section~\ref{sec:detailed-isa}.

\subsection{Normalization}

Using in silico data we observed that ISA has the best performance if the
input matrix is normalized (see \Rfunction{ISANormalize}). The
normalization produces two matrices: $E_r$ and
$E_c$. $E_r$ is calculated by transposing $E$ and
centering and scaling its genes (see the \Rfunction{scale} R function). $E_c$ is
calculated by centering and scaling the genes of $E$. $E_r$ is
used to calculate the sample signature of genes and $E_c$ is used
to calculate the signature of the samples.

It is possible to use another normalization, or not to use
normalization at all; the user has to construct an
\Rclass{ISAExpressionSet} object containing the three matrices
corresponding to the raw data, the gene-wise normalized data and the
sample-wise normalized data. This object can be passed to the
\Rfunction{ISAIterate} function. The matrices are not required to be
different, the user can supply the raw data matrix three times, if
desired.

\subsection{Gene and sample scores}

In addition to finding biclusters in the input matrix, the ISA also
assigns scores to the genes and samples, separately for each
module. The scores are between minus one and one and they are by
definition zero for the genes/samples that are not included in the
module. For the non-zero entries, the further the score of a
gene/samples is from zero, the stronger the association between the
gene/sample and the module. If the signs of two genes/samples are the
same, then they are correlated, if they have opposite signs, then they
are anti-correlated.

\section{Bicluster coherence and robustness measures}

\subsection{Coherence}

Madeira and Oliviera\cite{madeira04} define various coherence scores
for biclusters, these measure how well the rows and or columns are
correlated. It is possible to use these measures for ISA as well,
after converting the output of ISA to a \Rclass{biclust} object. 
We use the \texttt{Bc} object that was created in Section~\ref{sec:biclust}. 
Here are the measures for the first bicluster:
<<coherence>>=
constantVariance(exprs(ALL), Bc, number=1)
additiveVariance(exprs(ALL), Bc, number=1)
multiplicativeVariance(exprs(ALL), Bc, number=1)
signVariance(exprs(ALL), Bc, number=1)
@

You can use \Rfunction{sapply} to perform the calculation for many or
all modules, e.g. for this data set `constant variance' and `additive
variance' are not the same:
<<coherence all>>=
cv <- sapply(seq_len(Bc@Number), 
             function(x) constantVariance(exprs(ALL), Bc, number=x))
av <- sapply(seq_len(Bc@Number), 
             function(x) additiveVariance(exprs(ALL), Bc, number=x))
cor(av, cv)
@
Please see the manual pages of these functions and the paper cited
above for more details.

\subsection{Robustness}

The \Rpackage{eisa} package uses a measure that is related to
coherence; it is called robustness. Robustness is a generalization of
the singular value of a matrix. If there were no thresholding during
the ISA iteration, then ISA would be equivalent to a numerical method
for singular value decomposition and robustness would be the
same the principal singular value of the input matrix.

If the \Rfunction{isa} function was used to find the transcription
modules, then the robustness measure is used automatically to filter the results.
This is done by first scrambling the input matrix and then running ISA
on it. As ISA is an unsupervised algorithm it usually finds some
(although less and smaller) modules even in such a scrambled data
set. Then the robustness scores are calculated for the proper and the
scrambled modules and only (proper) modules that have a higher score
than the highest scrambled module are kept. The robustness scores are
stored in the seed data during this process, so you can check them
later:
<<robustness>>=
seedData(modules)$rob
@

\section{The \Rpackage{isa2} and \Rpackage{eisa} packages}%
\label{sec:isapackages}

ISA and its companion functions for visualization, functional
enrichment calculation, etc. are distributed in two separate R packages,
\Rpackage{isa2} and \Rpackage{eisa}.

\Rpackage{isa2} contains the implementation of ISA itself, and
\Rpackage{eisa} specifically deals with supplying expression data to
\Rpackage{isa2} and visualizing the results.

\section{Finer control over ISA parameters}%
\label{sec:detailed-isa}

The \Rfunction{ISA} function takes care of all steps performed during
a modular study, and for each step it uses parameters, that work reasonably
well. In some cases, however, one wants to access these steps
individually, to use custom parameters instead of the defaults.

In this section, we will still use the acute lymphoblastic leukemia
gene expression data from the \Rpackage{ALL} package.

\subsection{Non-specific filtering}

The first step of the analysis typically involves non-specific
filtering of the probesets. The aim is to eliminate the probesets that
do not show variation across the samples, as they only contribute
noise to the data.

By default (i.e. if the \Rfunction{ISA} function is called) this is
performed using the \Rpackage{genefilter} package, and the default
filter is based on the inter-quantile ratio of the probesets'
expression values, a robust measure of variance. Other possible
filtering techniques include using the AffyMetrix present/absent calls
produced by the \Rfunction{mas5calls} function of the \Rpackage{affy}
package, but this requires the raw data, so in this vignette we use a
simple method based on variance and minimum expression value: only
probesets that have a variance of at least \variable{varLimit} and
that have at least \variable{kLimit} samples with expression values
over \variable{ALimit} are kept.
<<filtering>>=
varLimit <- 0.5
kLimit <- 4
ALimit <- 5
flist <- filterfun(function(x) var(x)>varLimit, kOverA(kLimit,ALimit))
ALL.filt <- ALL[genefilter(ALL, flist), ]
@

The original expression set had \Sexpr{nrow(ALL)} features, the
filtered one has only \Sexpr{nrow(ALL.filt)}.

\subsection{Entrez Id matching}

In this step we match the probesets to Entrez Id and remove the ones
that don't map to any Entrez genes.
<<Entrez,cache=FALSE>>=
ann <- annotation(ALL.filt)
library(paste(ann, sep=".", "db"), character.only=TRUE)
ENTREZ <- get( paste(ann, sep="", "ENTREZID") )
EntrezIds <- mget(featureNames(ALL.filt), ENTREZ)
keep <- sapply(EntrezIds, function(x) length(x) >= 1 && !is.na(x))
ALL.filt.2 <- ALL.filt[keep,]
@

To reduce ambiguity in the interpretation of the results, we might
also want to keep only single probeset for each Entrez gene.
<<Entrez unique>>=
vari <- apply(exprs(ALL.filt.2), 1, var)
larg <- findLargest(featureNames(ALL.filt.2), vari, data=annotation(ALL.filt.2))
ALL.filt.3 <- ALL.filt.2[larg,]
@

\subsection{Normalizing the data}

The ISA works best, if the expression matrix is scaled and
centered. In fact, the two sub-steps of an ISA step require expression
matrices that are normalized differently. The
\Rfunction{ISANormalize} function can be used to calculate the
normalized expression matrices; it returns an \Rclass{ISAExpressionSet}
object. This is a subclass of \Rclass{ExpressionSet}, and contains
three expression matrices: the original raw expression, the row-wise
(=gene-wise) normalized and the column-wise (=sample-wise) normalized
expression matrix. The normalized expression matrices can be queried
with the \Rfunction{featExprs} and \Rfunction{sampExprs} functions.
<<normed>>=
ALL.normed <- ISANormalize(ALL.filt.3)
ls(assayData(ALL.normed))
dim(featExprs(ALL.normed))
dim(sampExprs(ALL.normed))
@

\subsection{Generating starting seeds for the ISA}

The ISA is an iterative algorithm that starts with a set of input
seeds. An input seed is basically a set of probesets and the ISA stepwise
refines this set by 1) including other probesets in the set that are
coexpressed with the input probesets and 2) removing probesets from it
that are not coexpressed with the rest of the input set.

The \Rfunction{generate.seeds} function generates a set of random
seeds (i.e. a set of random gene sets). See its documentation if you
need to change the sparsity of the seeds.

<<seeds>>=
set.seed(3)
random.seeds <- generate.seeds(length=nrow(ALL.normed), count=100)
@

In addition to random seeds, it is possible to start the ISA iteration
from ``educated'' seeds, i.e. gene sets the user is interested in, or
a set of samples that are supposed to have coexpressed genes. We
create another set of starting seeds here, based on the type of acute
lymphoblastic leukemia: ``B'', ``B1'', ``B2'', ``B3'', ``B4'' or
``T'', ``T1'', ``T2'', ``T3'' and ``T4''.
<<smart seeds>>=
type <- as.character(pData(ALL.normed)$BT)
ss1 <- ifelse(grepl("^B", type), -1, 1)
ss2 <- ifelse(grepl("^B1", type), 1, 0)
ss3 <- ifelse(grepl("^B2", type), 1, 0)
ss4 <- ifelse(grepl("^B3", type), 1, 0)
ss5 <- ifelse(grepl("^B4", type), 1, 0)
ss6 <- ifelse(grepl("^T1", type), 1, 0)
ss7 <- ifelse(grepl("^T2", type), 1, 0)
ss8 <- ifelse(grepl("^T3", type), 1, 0)
ss9 <- ifelse(grepl("^T4", type), 1, 0)
smart.seeds <- cbind(ss1, ss2, ss3, ss4, ss5, ss6, ss7, ss8, ss9)
@

The \variable{ss1} seed includes all samples, but their sign is
opposite for B-cell leukemia samples and T-cell samples. This way ISA
is looking for sets of genes that are differently regulated in these
two groups of samples. \variable{ss2} contains only B1 type samples,
so here we look for genes that are specific to this variant of the
disease. The other seeds are similar, for the other subtypes.

\subsection{Performing the ISA iteration}

We perform the ISA iterations for our two sets of seeds separately.
<<iteration>>=
modules1 <- ISAIterate(ALL.normed, feature.seeds=random.seeds, 
                        thr.feat=1.5, thr.samp=1.8, convergence="cor")
modules2 <- ISAIterate(ALL.normed, sample.seeds=smart.seeds,
                        thr.feat=1.5, thr.samp=1.8, convergence="cor")
@

\subsection{Dropping non-unique modules}

\Rfunction{ISAIterate} returns the same number of modules as the
number of input seeds; these are, however, not always meaningful,
the input seeds can converge to an all-zero
vector, or occasionally they may not converge at all. It is also
possible that two or more input seeds converge to the same module.

The \Rfunction{ISAUnique} function eliminates the all-zero or
non-convergent input seeds and keeps only one instance of the
duplicated ones.
<<unique>>=
modules1.unique <- ISAUnique(ALL.normed, modules1)
modules2.unique <- ISAUnique(ALL.normed, modules2)
@

\Sexpr{length(modules1.unique)} modules were kept for the first set of
seeds and \Sexpr{length(modules2.unique)} for the second set.

\subsection{Dropping non-robust modules}

The \Rfunction{ISAFilterRobust} function filters a set of modules by
running ISA with the same parameters on the scrambled data set and then
calculating a robustness score, both for the real modules and the ones
from the scrambled data. The highest robustness score obtained from
the scrambled data is used as a threshold to filter the real modules.
<<robust>>=
modules1.robust <- ISAFilterRobust(ALL.normed, modules1.unique)
modules2.robust <- ISAFilterRobust(ALL.normed, modules2.unique)
@

We still have \Sexpr{length(modules1.robust)} modules for
the first set of seeds and \Sexpr{length(modules2.unique)} for the
second set.

\subsection{Differentially regulated modules}

Now we check whether the ISA modules that we found can be used as
classifiers for the different types of ALL. For this we use the sample
scores of the modules. The score of a sample is the weighted average
of the normalized expression level of the modules genes. Thus, a good
classifier module should have high sample scores for one type of ALL
and low for the other. Here we perform a $t$-test to quantify this
difference.

Let's first check, whether any of the modules can distinguish between
T-cell and B-cell ALL samples. We perform $t$-tests for both sets of
modules.
<<B vs T>>=
scores1 <- getSampleMatrix(modules1.robust)
tt1 <- colttests(scores1, as.factor(substr(type, 1, 1)))
scores2 <- getSampleMatrix(modules2.robust)
tt2 <- colttests(scores2, as.factor(substr(type, 1, 1)))
sign1 <- which(p.adjust(tt1$p.value, "holm") < 0.05)
sign2 <- which(p.adjust(tt2$p.value, "holm") < 0.05)
@

For the first (random) set of seeds \Sexpr{length(sign1)} modules show
significant difference for T-cell and B-cell samples, for the second
(smart) set \Sexpr{length(sign2)} of them.

Let's make some condition plots for the best separating modules from
each set (Fig.~\ref{fig:dysregmodules}).
<<conditions-plots-dysreg,results=hide,include=FALSE,eval=FALSE>>=
color <- ifelse(grepl("T", type), "orange", "darkolivegreen")
layout(cbind(1:2))
condPlot(modules1.robust, which.min(tt1$p.value), ALL.normed, col=color, 
          main="Best separator, random seeds")
condPlot(modules2.robust, which.min(tt2$p.value), ALL.normed, col=color,
          main="Best separator, smart seeds")
@

\begin{figure}
<<conditions-plots-dysreg-real,fig=TRUE,width=10,height=8,echo=FALSE,results=hide>>=
<<conditions-plots-dysreg>>
@
\caption{Condition plots for the best separator module found with
  random seeds and the one found with the smart seeds. They are
  essentially the same, with some minor differences. 
}
\label{fig:dysregmodules}
\end{figure}

Let's extract the modules that are good separators.
<<extract separators>>=
modules1.TB <- modules1.robust[[sign1]]
modules2.TB <- modules2.robust[[sign2]]
@

As it turns out, one modules found using the smart seeds was also
found with random seeds.

\subsection{Enrichment calculations}

We can check our separator modules against the Gene Ontology
categories and the pathways in the KEGG database to find dysregulated
GO categories and/or KEGG pathways.
<<dysreg enrichment>>=
GO.dysreg1 <- ISAGO(modules1.TB)
GO.dysreg2 <- ISAGO(modules2.TB)
KEGG.dysreg1 <- ISAKEGG(modules1.TB)
KEGG.dysreg2 <- ISAKEGG(modules2.TB)
@

Let's collect the significantly enriched GO categories and KEGG
pathways.
<<dysreg enriched categories>>=
gocats <- unique(unlist(lapply(c(GO.dysreg1, GO.dysreg2), function(x) 
                               unique(unlist(sigCategories(x))))))
keggp  <- unique(unlist(lapply(list(KEGG.dysreg1, KEGG.dysreg2), function(x)
                        unique(unlist(sigCategories(x))))))
sapply(mget(gocats, GOTERM), Term)
mget(keggp, KEGGPATHID2NAME)
@

\subsection{More separator modules}

Of course we can search for modules that can separate the samples of
the ALL subtypes as well, e.g. let's try to find some that
differentiate between type B1 and other B types. 
<<B1 vs others>>=
keep <- grepl("^B[1234]", type)
type.B <- ifelse(type[keep]=="B1", "B1", "Bx")
scores.B1.1 <- scores1[keep,]
scores.B1.2 <- scores2[keep,]
tt.B1.1 <- colttests(scores.B1.1, as.factor(type.B))
tt.B1.2 <- colttests(scores.B1.2, as.factor(type.B))
min(p.adjust(na.omit(tt.B1.1$p.value)))
min(p.adjust(na.omit(tt.B1.2$p.value)))
@

Both module sets seem to have some good separators, let us make some
condition plots (Fig.~\ref{fig:B1condplots}).
<<B1-cond-plots,results=hide,include=FALSE,eval=FALSE>>=
color <- ifelse(type == "B1", "green", 
                ifelse(grepl("^T", type), "orange", "darkolivegreen"))
layout(cbind(1:2))
condPlot(modules1.robust, which.min(tt.B1.1$p.value), ALL.normed, col=color, 
          main="Best B1 separator, random seeds")
condPlot(modules2.robust, which.min(tt.B1.2$p.value), ALL.normed, col=color,
          main="Best B1 separator, smart seeds")
@

\begin{figure}
\centering
<<B1-cond-plots-real,fig=TRUE,echo=FALSE,results=hide,width=10,height=8>>=
<<B1-cond-plots>>
@ 
\caption{Condition plots for the modules that best separate the B1-type 
  ALL samples from the other B-cell ALL samples. The two modules are essentially
  the same.}
\label{fig:B1condplots}
\end{figure}

From the plot it seems that these two modules are essentially the
same, one is the opposite of the others.
<<check B1 same module>>=
B1.cor <- c(cor(scores1[,which.min(tt.B1.1$p.value)],
                scores2[,which.min(tt.B1.2$p.value)]),
            cor(getFeatureMatrix(modules1.robust, 
                                 mods=which.min(tt.B1.1$p.value)),
                getFeatureMatrix(modules2.robust, 
                                 mods=which.min(tt.B1.2$p.value))))
B1.cor
@

Indeed, they are almost the same, their Pearson correlation is around
\Sexpr{round(B1.cor[1],3)} for the sample scores and
\Sexpr{round(B1.cor[2],3)} for the feature (=gene) scores.

\section{More information}

For more information about the ISA, please see the references
below. The ISA homepage at
\url{http://www.unil.ch/cbg/homepage/software.html} has example data
sets, and all ISA related tutorials and papers.

\section{Session information}

The version number of R and packages loaded for generating this
vignette were:

<<sessioninfo,results=tex,echo=FALSE>>=
toLatex(sessionInfo())
@

\bibliographystyle{apalike}
\bibliography{EISA}
\nocite{*}

\end{document}