### 2010-01-14 expressionviewnote/expressionview.tex
Added Sven's changes to ExpressionView note.

\documentclass[round]{bioinfo}

\usepackage{color}
\newcommand{\cred}{\color{red}}

\usepackage{url}
% \usepackage[colorlinks,citecolor=blue,linkcolor=red,urlcolor=blue]{hyperref}
\newcommand{\thefirstpage}{1} \newcommand{\thelastpage}{1}

\copyrightyear{2010}
\pubyear{2010}

\begin{document}

\firstpage{1}
\application{}
\title[ExpressionView]{ExpressionView - an interactive viewer for
  modules identified in gene expression data} 
\author[Andreas L\"uscher, G\'abor Cs\'ardi, Aitana Morton de
Lachapelle, Zolt\'an Kutalik, and Sven Bergmann]{Andreas
  L\"uscher$^{1\dagger}$, G\'abor Cs\'ardi$^{1,2\dagger}$, Aitana Morton de
  Lachapelle$^{1,2\dagger}$, Zolt\'an Kutalik$^{1,2\dagger}$, 
  Bastian Peter$^{1,2}$
  and Sven Bergmann$^{1,2}$\footnote{to whom correspondence should be
    addressed}}  
\address{
  $^{1}$Swiss Institute of Bioinformatics, Lausanne, Switzerland\\
  $^{2}$Department of Medical Genetics, University of Lausanne,
  Lausanne, Switzerland\\
  +: equal contribution
}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\maketitle

\begin{abstract}

\section{Summary:}
ExpressionView is an R package that provides an interactive graphical
environment to explore transcription modules identified in gene expression
data. A sophisticated ordering algorithm is used to present the
modules with the expression in a visually appealing layout that provides an intuitive
summary of the results. From this overview, the user can select
individual modules and access all the biologically relevant metadata
associated with it.

\section{Availability:} 
%\href{http://www.unil.ch/cbg/ExpressionView}{R 
%  package with open source license,
\url{http://www.unil.ch/cbg/ExpressionView}
%}

\section{Contact:} \href{Sven.Bergmann@unil.ch}{Sven.Bergmann@unil.ch}

\section{Supplementary information:}
Screenshots, tutorials and sample data sets can be found on the
ExpressionView \href{http://www.unil.ch/cbg/ExpressionView}{website}. 
\end{abstract}

\vspace*{-10pt}
\section{Introduction}
Biclustering is an unsupervised data analysis method which is
frequently used to explore microarray data. Biclustering algorithms
identify co-expressed groups of
genes and conditions (samples) where this co-expression occurs from
collections of expression profiles. We refer to such groups (or
biclusters) as modules. While
there is a multitude of biclustering software available~(see in
\cite{madeira04,ihmels04}), packages with intuitive interfaces that
allow for an interactive exploration of the results are
sparse~\citep{santamaria08}. ExpressionView is designed to close this
gap, and facilitate the collaboration between bioinformaticians and
life scientists without consolidated programming
experience. Implemented as an R package, the user can apply all the
powerful microarray analysis tools provided by the
Bioconductor~\citep{gentleman04} project before exporting the data to
a platform independent Adobe Flash applet. This stand-alone applet
provides and integrated, intuitive and interactive visualization of
the modules together with the underlying gene expression data and
additional information on the genes and conditions.

\section{Package design and workflow}
The ExpressionView package contains two independent parts: An ordering
algorithm implemented in C++ and an Adobe Flash applet written in
ActionScript and Adobe Flex. Fig.~\ref{fig:workflow} schematically
summarises the typical workflow that uses the gene expression data
available as a Bioconductor ExpressionSet and combines it with the
biclustering results to produce an ExpressionView data file that can
be explored with the Flash applet. In the following, we describe the
different steps in more detail. 
\begin{figure*}[!tpb]
\centerline{\includegraphics[width=0.7\linewidth]{fig1-crop}}
\caption{How to use ExpressionView: Starting from gene expression data
  in the form of a Bioconductor ExpressionSet, the user first runs a
  biclustering algorithm to determine co-expressed groups of genes and
  the relevant samples. In a second step, the rows and columns of the gene
  expression matrix are rearranged to produce an easily readable
  overview of the results. The last step consists of combining the
  gene expression data and its associated metadata with the results
  from the biclustering and produce an ExpressionView data file that
  can be explored with the interactive Flash
  applet.}\label{fig:workflow} 
\end{figure*}
For users who prefer to install the viewer as a standalone program, we
also provide a platform independent Adobe AIR version that can be
downloaded from the
\href{http://www.unil.ch/cbg/ExpressionView}{website}.

\subsection{Gene expression data and modules}
ExpressionView is designed to work with gene expression data in the
form of a Bioconductor ExpressionSet. This class provides a
user-friendly way to access the actual gene expression matrix and its
associated metadata. There is a variety of biclustering algorithms
described in the literature~\citep{madeira04,prelic06}, several of
which have been implemented as R packages. ExpressionView can treat
results obtained by the Iterative Signature Algorithm
(ISA)~\citep{bergmann03} (see \citep{csardi09} for its R
implementation) and any of the methods available in the Biclust
package~\citep{kaiser08}. Since the structure of biclustering results
is independent of the algorithm, an extension to other methods is
straightforward.

\subsection{Reordering genes and conditions}
To present the collection of possibly overlapping modules in a visually
appealing form, it is necessary to reorder the rows (genes) and
columns (conditions) of the
gene expression matrix in such a way that biclusters form contiguous
rectangles. Since it is in general impossible to find such an
arrangement for more than two mutually overlapping modules, we propose
here an approximate solution that optimizes the arrangement within the
original data by maximising the total area of the largest contiguous
module subsets. (An alternative would be to repeat rows and columns as
necessary~\cite{grothaus06}, but for many modules this will results in
a very large expression matrix.)

This optimization task is an interesting problem on its own, which to
the best of our knowledge has not been studied in the literature. Thus
we briefly outline our strategy here (see our website for details). 
The reordering of the rows and the columns do not depend on each
other, so the same optimization method can be applied sequentially,
say first to rows and then to columns. For a given order of the
elements (either genes or conditions) we compute for each module the
size of the largest contiguous sequence of elements (i.e. the maximal
number of neighboring elements $N^{max}_i$). Then as a measure of the
quality $Q$ of the order we sum this quantity over all modules
($Q=\sum_i N^{max}_i$). The optimization procedure for $Q$ starts from
some initial order. Two operations are then applied to this order: (1)
Permutations that exchange two elements within a module and (2) shifts
of a sequence of elements within a module to a different position. We
use a greedy iterative scheme that performs these operations at
randomly selected positions and keeps the new order if it improves
$Q$. The algorithm stops if after a given number of operations no
significant improvement of $Q$ is achieved.

To determine the efficiency of this method, we have studied a large
number of perfectly orderable, but initially scrambled, test cases. We
find that in almost every situation, the proposed algorithm finds an
order that recovers more than 99\% of the score of the optimal
solution and in most cases, it recovers the correct
alignment. For random samples, which are more representative for
actual gene expression data, the execution time increases polynomially
with the number of clusters $m$ as ${\mathcal O}(m^\alpha)$, where
$\alpha \in [1.6, 2]$, almost independent of the number of elements
$n$. For a given number of clusters, we find ${\mathcal O}(n^\alpha)$,
with $\alpha \in [2.5, 2.7]$.

\subsection{Export}
Once the optimal order is determined, the program rearranges the gene
expression matrix accordingly and exports all the relevant information
into an XML file. This format has the advantage of being
self-explanatory and extensible. In addition, the files can be read
and edited with any text editor. The actual gene expression data is
rounded to two digit precision and encoded in Base64. The structure of
the XML file is described in more detail on the ExpressionView
\href{http://www.unil.ch/cbg/ExpressionView}{website}. In most cases,
these files are of the order of a few Megabytes and can thus easily be
sent by e-mail.

\subsection{Visualisation}
The core application of this package is the Flash applet that allows
the user to interactively explore the gene expression data and the
modules. The Flash architecture has the advantage of
producing a platform and browser independent applet that can be
launched from any computer with even the most restrictive user
rights. A screenshot is shown in Fig.~\ref{fig:workflow}. The
interface is divided in two parts: On the left-hand side, the user
finds the gene expression data in the common heat map form, on top of
which the modules are overlaid. On the right-hand side, the metadata
associated with the gene expression data and the results of the
enrichment calculations, GO (Gene Ontology)~\citep{ashburner00} and
KEGG (Kyoto Encyclopedia of Genes and Genomes)~\citep{kanehisa04}, are
shown. Wherever possible, these elements are linked to the
corresponding databases. The interface essentially behaves as an image
viewer, allowing the user to zoom and pan around the expression data,
getting instant feedback on the selected item. The currently selected
region can be exported to a PDF file, helpful to prepare figures for
presentations or articles, or a simple text file.

\section*{Acknowledgement}

\paragraph{Funding\textcolon} The authors are grateful to the Swiss
Institute of Bioinformatics, the Swiss National Science Foundation
(3100AO-116323/1) and the European Framework Project 6 (through
the EuroDia and AnEuploidy projects).

\paragraph{Conflict of interest\textcolon} none declared.

\begin{thebibliography}{6}

\bibitem[Ashburner et~al.(2000)]{ashburner00}
Ashburner,~M. \emph{et al.} (2000)
Gene Ontology: tool for the unification of biology.
\href{http://dx.doi.org/10.1038/75556}{\emph{Nature Genetics}, {\bf 25}, 25--29.}

\bibitem[Bergmann et~al.(2003)]{bergmann03}
Bergmann,~S., Ihmels,~J. and Barkai,~N. (2003)
Iterative signature algorithm for the analysis of large-scale gene expression data.
\href{http://dx.doi.org/10.1103/PhysRevE.67.031902 }{\emph{Physical Review E}, {\bf 67}, 031902.}

%\bibitem[Cs{\'a}rdi et~al.(2009)]{csardi09}
%Cs{\'a}rdi,G., Kutalik,Z. and Bergmann,S. (2009)
%Modular analysis of gene expression data with R.
%\href{http://dx.doi.org/}{\emph{Bioinformatics}, ...}

\bibitem[Cs{\'a}rdi(2009)]{csardi09}
G\'abor Cs\'ardi (2009)
isa2: The Iterative Signature Algorithm. 
{R package version 0.2.}

\bibitem[Gentleman et~al.(2004)]{gentleman04}
Gentleman,~R.C., \emph{et al.} (2004)
Bioconductor: open software development for computational biology and bioinformatics.
\href{http://dx.doi.org/10.1186/gb-2004-5-10-r80}{\emph{Genome Biology}, {\bf 5}, R80.}

\bibitem[Grothaus et~al.(2006)]{grothaus06}
Grothaus,~G., Mufti,~A., and Murali,~TM. (2006)
Automatic layout and visualization of biclusters.
\href{http://dx.doi.org/10.1186/1748-7188-1-15}{\emph{Algorithms for Molecular Biology}, {\bf 1}, 15.}

\bibitem[Ihmels and Bergmann(2004)]{ihmels04}
Ihmels JH and Bergmann S. Challenges and prospects in the analysis
of large-scale gene expression data. (2004)
{\emph{Brief Bioinform} \bf{5} 313--327.}

\bibitem[Kaiser and Leisch(2008)]{kaiser08}
Kaiser,~S. and Leisch,~F. (2008)
A toolbox for bicluster analysis in R.
\href{http://epub.ub.uni-muenchen.de/3293/}{Technical Report 028, Department of Statistics, University of Munich}.

\bibitem[Kanehisa et~al.(2004)]{kanehisa04}
Kanehisa,~M. \emph{et al.} (2004)
The KEGG resource for deciphering the genome.
\href{http://dx.doi.org/10.1093/nar/gkh063}{\emph{Nucleic Acids Res.}, {\bf 32} (Database special issue), D277.}

\bibitem[Madeira and Oliveira(2004)]{madeira04}
Madeira,~S.C. and Oliveira,~A.L. (2004)
Biclustering algorithms for biological data analysis: A survey.
\href{http://dx.doi.org/10.1109/TCBB.2004.2}{\emph{IEEE/ACM Transactions on Computational Biology and Bioinformatics}, {\bf 1}, 24--45.}

\bibitem[Prelic et~al.(2006)]{prelic06}
Prelic,~A., \emph{et~al.} 
A systematic comparison and evaluation of biclustering methods for gene expression data 
\href{http://dx.doi.org/10.1093/bioinformatics/bt1060}{\emph{Bioinformatics}, {\bf 22}(9), 1122--1129.}

\bibitem[Santamar\'ia et~al.(2008)]{santamaria08}
Santamar\'ia,~R., Ther\'on,~R. and Quintales,~L. (2008)
BicOverlapper: A tool for bicluster visualization. 
\href{http://dx.doi.org/10.1093/bioinformatics/btn076}{\emph{Bioinformatics}, {\bf 24}, 1212--1213.}

\end{thebibliography}

\end{document}