### 2010-02-04 isanote/answers_to_reviewers.tex
ISA application note: added 'response to reviewers' skeleton.

\documentclass[a4paper]{article}
\usepackage{url}

\title{Modular analysis of gene expression data with R\\
  Answers to reviewers}
\author{G\'abor Cs\'ardi, Zolt\'an Kutalik and Sven Bergmann}

\begin{document}

\maketitle

\section{Reviewer 1}

\begin{quote}
Comments to the Author

This Application Note presents a software package that implements the
Iterative Signature Algorithm and provides a convenient visualization
of it’s results. The method is useful and the software may be of help
to interested users. 
\end{quote}

\begin{quote}
I have two problems with the Note in its present form.
The first may be due to my misunderstanding of the method. My
understanding was that some threshold Tg is applied on the genes and
another one, Tc, on the conditions. The aim is to reach a bicluster (a
set of conditions C and a set of genes G) such that G contains only
genes with expression above Tg for all the conditions of C, and
conditions are included in C only if the expression of all genes of G
exceeds Tc (one can replace a lower threshold with an upper one). The
heatmap of the Figure implies that my understanding was incorrect –
the conditions satisfy a double threshold (expressions are either
above or below some thresholds) and the genes have expressions above a
Tg in some conditions or below some other Tg’ in other conditions. 
I urge the authors to define what specifies or defines a bicluster –
do it briefly and sacrifice a few lines from elsewhere. 
\end{quote}

\begin{quote}
My second concern is minor – the abstract implies that the ISA is
special in allowing overlapping cluster assignments. This is incorrect
– there are other biclustering methods that do this. If I am not
mistaken, all three original biclustering methods that came out in
2000 (Cheng, Getz and Califano) allow this. 
\end{quote}

\section{Reviewer 2}

\begin{quote}
Comments to the Author

Summary

This manuscript describes two software packages for GNU R, for
discovering biclusters (here called transcription modules. see [1] for
a survey) in a given gene expression matrix based on the Iterative
Signature Algorithm (ISA) [2]. The first package, called isa2,
contains the implementation of the basic ISA, and can be used to
analyze any tabular data. The second package, called eisa, is built on
isa2. It adds support to standard BioConductor [3] data structures,
and contains gene expression specific tools for visualization of
biclusters. Both packages support execution of ISA with default
parameters, and also execution of ISA step by step with user-specified
parameters. The steps of ISA are normalization, iteration, filtering
of modules and enrichment calculation. In the current note, the
authors present both packages, and describe how to install and run
them step by step, using the acute lymphoblastic leukemia data set
(ALL) [4] as an example. 
\end{quote}

\begin{quote}
Overall evaluation

The manuscript and its accompanying tutorials are well written and
give an adequate theoretical background to ISA. The tutorials include
many detailed and visualized examples that make learning the package
quite easy for an R user. However, the following major points should
be taken into consideration: 
\end{quote}

\begin{quote}
1.      The manuscript and the website lack explanations about how the
authors implemented ISA and how they used the advantages of R. The
authors emphasize that the packages are highly optimized and run
extremely fast even for very large datasets.  Therefore, the
accompanying material should contain a summary describing how this
efficiency was achieved. In addition, a graph showing execution time
as a function of dataset size should be given. 
\end{quote}

\begin{quote}
2.      The tutorials give examples of execution on a single rather
old dataset (ALL) which is quite small in comparison to datasets of
recent years. Here the authors should give results for ISA execution
on more up-to-date datasets, and on heterogeneous datasets containing
profiles from multiple studies. 
\end{quote}

\begin{quote}
To summarize, the software is well motivated and enables users of the
R environment to utilize ISA. I suggest accepting the manuscript
conditional on solving the issues mentioned in this report. 
\end{quote}

\begin{quote}
Additional important remarks:

1.      The manuscript should contain references to additional
biclustering implementations, in R or in other environments, so that
the user could choose the one that best fits his needs in terms of
input/output, species, etc. Examples are SAMBA (via the Expander
software) [5,6] and BicAt [7]. 
\end{quote}

\begin{quote}
2.      The sample score plots example given in the EISA tutorial
(\url{http://www2.unil.ch/cbg/index.php?title=EISA\_tutorial}, section 6.5)
is not sufficiently clear. What are the scores represented by the
horizontal lines and how can one calculate them? Moreover, how can one
separate B-cell and T-cell leukemia samples using this plot? 
\end{quote}

\begin{quote}
3.      It is not quite clear what is the difference between the ISA
implementation included in the isa2 package and the ISA implementation
included in the eisa package. 
\end{quote}

\begin{quote}
4.      It should be clarified who are the potential users for the ISA
implementation for R environment. Assuming these are mainly
biologists, bioinformaticians and biostatisticians (and not only
programmers), more technical details of how to work with the packages
should be given. The following points are important examples: 
a.      Since the ISA implementation in eisa package imports data from
another package (e.g. hgu95av2.db), a tutorial for how to create such
a .db package from a given expression data text file should be given
(or referred to). 
\end{quote}

\begin{quote}
b.      Since both ISA packages use various data structures, a short
explanation of how to construct and use them should be given (or
references to such explanation), and also how to export data from
these structures to external files. 
\end{quote}

\begin{quote}
Minor remarks:

1.      The authors describe methods that perform enrichment tests for
the gene sets corresponding to the ISA modules against various
databases. In particular, they perform enrichment analysis against GO
database, KEGG pathway database, chromosomes and the TargetScan [8]
database. One may also want to investigate the correspondence of
biclusters and protein-protein interaction networks (which have been
derived from other types of data than gene expression data), e.g. the
DIP database [9]. An example for such analysis can be found in [10]. 
\end{quote}

\begin{quote}
2.      The heatmap image of a transcription module obtained by any
one of the two methods presented in the tutorial should include a
color legend. 
\end{quote}

\begin{quote}
3.      Since ISA function accepts an ExpressionSet, one must obtain
annotations for its expression data file in order to run ISA function
on it. Annotations should not be a prerequisite for running ISA. 
\end{quote}

\begin{quote}
4.      In section 10.8 it is mentioned that two significant modules
were found using random seeds and two significant (separating) modules
were found using smart seeds. However, in figure 7 it is noted that
only one module found using smart seeds. At the end of the section it
is mistakenly written "As it turns out, one modules found". 
\end{quote}

\begin{quote}
References:
[1] A. Tanay, R. Sharan, and R. Shamir. Biclustering algorithms: A
survey. In E. by Srinivas Aluru, editor, In Handbook of Computational
Molecular Biology. Chapman \& Hall/CRC, Computer and Information
Science Series, 2005.

[2] Bergmann, S., Ihmels, J., and Barkai, N. (2003). Iterative
signature algorithm for the analysis of large-scale gene expression
data. Phys Rev E Nonlin Soft Matter Phys, page 031902.

[3] Gentleman, R. C., Carey, V. J., Bates, D. M., et
al. (2004). Bioconductor: Open software development for computational
biology and bioinformatics. Genome Biology, 5, R80.

[4] Chiaretti, S., Li, X., Gentleman, R., Vitale, A., Vignetti, M.,
Mandelli, F., Ritz, J., and Foa, R. (2004). Gene expression profile of
adult t-cell acute lymphocytic leukemia identifies distinct subsets of
patients with different response to therapy and survival. Blood,
103(7).

[5] R. Sharan, A. Maron-Katz, N. Arbili, and R. Shamir.EXPANDER:
EXPression ANalyzer and DisplayER, 2002.Software package, Tel-Aviv
University, \url{http://www.cs.tau.ac.il/~rshamir/expander/expander.html}.

[6] A. Tanay, R. Sharan, M. Kupiec, and R. Shamir. Revealing
modularity and organization in the yeast molecular network by
integrated analysis of highly heterogeneous genomewide data. Proc Natl
Acad Sci U S A., 101(9):2981–6, 2004.

[7] Barkow, S., et al. (2006) BicAT: a biclustering analysis
toolbox. Bioinformatics, 22, 1282–1283

[8] Targetscan [\url{http://www.targetscan.org/}]

[9] Salwinski, L., et al. (2004) The Database of Interacting Proteins:
2004 update. Nucleic Acids Res, . 32, D449–D451

[10] Prelic,A. et al. (2006) A systematic comparison and evaluation of
biclustering methods for gene expression data. Bioinformatics, 22,
1122–1129. 
\end{quote}

\end{document}