### 2010-04-21 RExpressionView/inst/doc/ExpressionView.tutorial.Rnw
EV: reindented vignettes.

% \VignetteIndexEntry{ExpressionView}
\documentclass{article}
\usepackage{ragged2e}
\usepackage{hyperref}
\usepackage{url}
\usepackage[margin=3cm]{geometry}

\newcommand{\Rfunction}[1]{\texttt{#1}}
\newcommand{\Rpackage}[1]{\texttt{#1}}
\newcommand{\Rclass}[1]{\texttt{#1}}
\newcommand{\Rargument}[1]{\textsl{#1}}
\newcommand{\filename}[1]{\texttt{#1}}
\newcommand{\variable}[1]{\texttt{#1}}

%\SweaveOpts{prefix.string=graphics/plot}

\setlength{\parindent}{2em}

\begin{document}

\title{ExpressionView}
\author{Andreas L\"uscher}
\maketitle

\tableofcontents

\RaggedRight

<<set width,echo=FALSE,print=FALSE>>=
options(width=60)
options(continue=" ")
@

\section{Introduction}
Clustering genes according to their expression profiles is an
important task in analyzing microarray data. In this tutorial, we
explain how to use ExpressionView, an R package designed to
interactively explore biclusters identified in gene expression data,
in conjunction with the Iterative Signature Algorithm
(ISA)~\cite{bergmann03} and the biclustering methods available in the
\Rpackage{Biclust} package~\cite{kaiser08}.

\section{Loading the gene expression data}
The \Rpackage{ExpressionView} package requires the gene expression
data to be available in the form of a BioConductor
\Rclass{ExpressionSet}. In this tutorial we will use the BioConductor
sample data from a clinical trial in acute lymphoblastic leukemia
provided by the \Rpackage{ALL} package.
<<load the data>>=
library(ALL)
library(hgu95av2.db)
data(ALL)
@
The data set contains \Sexpr{ncol(ALL)} samples and \Sexpr{nrow(ALL)}
features.

\section{Find biclusters}
There are many biclustering algorithms described in the
literature~\cite{madeira04}. All of them aim to reduce the complexity
of the gene expression data by identifying suitable groups of genes
and conditions that are co-expressed. In this tutorial show how to use
\Rpackage{ExpressionView} with some of the available biclustering
algorithms.

\subsection{Iterative Signature Algorithm (ISA)}
The ISA~\cite{bergmann03} for gene expression data is implemented in
the \Rpackage{eisa} package:
<<load the packages>>=
library(eisa)
@

To run the ISA for the given data set, simply call the \Rfunction{ISA}
function on the
\Rclass{ExpressionSet} object:
<<ISA, eval=FALSE>>=
set.seed(5) # initialize random number generator to get always the same results
modules <- ISA(ALL)
@ 
Depending on your computing resources, this should take roughly two
minutes. If you do not want to wait that long, you can shorten the
calculation by selecting the thresholds for genes and conditions:
<<fast ISA, cache=FALSE>>=
threshold.genes <- 2.7 
threshold.conditions <- 1.4 
set.seed(5)
modules <- ISA(ALL, thr.gene=threshold.genes, thr.cond=threshold.conditions)
@
If you leave the thresholds undefined, as in the first example, the
ISA runs with the default values, i.e.,
\variable{thr.gene=c(2,2.5,3,3.5,4)} and
\variable{thr.cond=c(1,1.5,2,2.5,3)}. In both cases, the random number
generator is initialized manually using \Rfunction{set.seed(5)}, to
give reproducible results. The \Rfunction{isa} function returns an
\Rclass{ISAModules} object. Typing its name returns a brief summary of
the results:
<<ISAModules summary>>=
modules
@

This object can be directly used with the functions of the
\Rpackage{ExpressionView} package. See Section~\ref{sec:ordering} for
details.

\subsection{Algorithms of the \Rpackage{Biclust} package}
The \Rpackage{biclust} package implements several biclustering
algorithms in a unified framework. It uses the \Rclass{Biclust} class 
to store a set of biclusters. Let us use the Plaid Model Bicluster
Algorithm~\cite{turner04} on the \Rpackage{ALL} data set
<<biclust bcplaid,cache=FALSE>>=
library(biclust)
biclusters <- biclust(exprs(ALL), BCPlaid(), fit.model=~m+a+b, verbose=FALSE)
@

\Rclass{Biclust} objects can be directly used with the
\Rpackage{ExpressionView} functions. Alternatively, they can be
converted to \Rclass{ISAModules} objects, using the standard
\Rfunction{as} R function:
<<isa>>=
as(biclusters, "ISAModules")
@
results an \Rclass{ISAModules} object.

\subsection{External clustering programs}
Since the structure of biclustering results is independent of the
applied method, it is straightforward to import results obtained from
external clustering programs and convert them to
\Rclass{ISAModules}. To illustrate the conversion, let us consider the
sample data and {\bf randomly} assign the  \Sexpr{nrow(ALL)} genes and
\Sexpr{ncol(ALL)} samples to  \Sexpr{length(modules)} modules. The
resulting modules can be described by two binary matrices
<<random-modules>>=
modules.genes <- matrix(as.integer(runif(nrow(ALL) * length(modules)) > 0.8), 
                        nrow=nrow(ALL))
modules.conditions <- matrix(as.integer(runif(ncol(ALL) *
                                        length(modules))>0.8),
                        nrow=ncol(ALL))
@
indicating if a given gene \variable{i} is contained in module
\variable{j} if \variable{modules.genes[i,j]$\ne$0}. Using these
matrices, it is straightforward to create an \Rclass{ISAModules}
object: 
<<toisa>>=
new("ISAModules",
    genes=modules.genes, conditions=modules.conditions,
    rundata=data.frame(), seeddata=data.frame())
@

\section{Order}%
\label{sec:ordering}
To present the tens of possibly overlapping biclusters in a visually
appealing form, it is necessary to reorder the rows and columns of the
gene expression matrix in such a way that biclusters form contiguous
rectangles. Since for more than two mutually overlapping biclusters,
it is in general impossible to find such an arrangement, one has to
make concessions. In contrast methods that propose to repeat rows and
columns as necessary to achieve this goal~\cite{grothaus06}, we prefer
to optimize the arrangement within the original data by maximizing the
area of the largest contiguous biclusters.

The \Rfunction{OrderEV} function implemented in the
\Rpackage{ExpressionView} package determines the optimal order of the
gene expression matrix for a given set of biclusters. It can be called
with \Rclass{ISAModules} or \Rclass{Biclust} objects as the first argument:
<<expressionview, results=HIDE>>=
library(ExpressionView)
optimalorder <- OrderEV(modules)
@

The result is a list containing various mappings between the original
data and the optimal arrangement. Note that the genes and the samples
can be ordered separately. Apart form reordering the full gene
expression matrix, the algorithm also determines the best arrangement
of individual biclusters. The mapping of the genes and the samples
contained in bicluster \variable{i} can be accessed by
<<gene maps, eval=FALSE>>=
optimalorder$genes[i+1]
optimalorder$samples[i+1]
@
The first elements of the lists contain the optimal ordering of the
complete matrix. By default, the \Rfunction{OrderEV} function
runs for roughly one minute, this might not be sufficient to find an
appropriate order for data containing many overlapping biclusters. The
status of the ordering is stored in
<<status ordering>>=
optimalorder$status
@
% $
If the status is set to \variable{1}, the algorithm has found the
optimal solution. A \variable{0} indicates that the the calculation
could not be terminated within the given timeframe. The
\Rfunction{OrderEV} function accepts two additional parameters to
circumvent the problem of partial alignment: One can start the
ordering from a given initial configuration, i.e., the result of a
previous arrangement by defining the \Rargument{initialorder} argument
<<status ordering, eval=FALSE>>=
optimalorderp <- OrderEV(modules, initialorder=optimalorder, maxtime=120)
@
and one can increase the time limit by specifying
\Rargument{maxtime}. Note that the time is indicated in seconds and
cannot be  smaller than 1.

\section{Export}
The \Rfunction{ExportEV} function allows the user to combine the
available data and export it to an XML file that can be read by the
Flash applet:
<<export, eval=FALSE>>=
ExportEV(modules, ALL, optimalorder, filename="file.evf")
@
The function gathers the data contained in the \Rclass{ExpressionSet}
\variable{ALL}, orders it according to the optimal arrangement
\variable{optimalorder} and adds the biclusters defined in
\variable{modules}. The output is an uncompressed XML file that can be
opened with any text viewer. We have chosen to use the extension
\variable{.evf} (for ExpressionView file) for the data files. This
extension is associated with the stand-alone version of the viewer, so
that one can simply double-click on such a file to launch the program
and load the data. The file association is the reason why we do not
use the \variable{.xml} extension. A description of the XML layout can
be found on the \Rpackage{ExpressionView} website at
\url{http://www.unil.ch/cbg/ExpressionView}. Before exporting the
data, the \Rfunction{ExportEV} function automatically calculates
GO~\cite{ashburner00} and KEGG~\cite{kanehisa04} enrichments for the
given biclusters.

\section{Visualize}
The ExpressionView Flash applet can be launched from the R
environment:
<<export, eval=FALSE>>=
LaunchEV()
@
Video tutorials describing how to use the applet can be found on the
ExpressionView website at
\url{http://www.unil.ch/cbg/ExpressionView}. The screenshot shown in
Fig.~\ref{fig:screenshot} and the description below illustrate the
main features of the applet:
\begin{figure} 
  \begin{center} 
    \includegraphics[width=0.95\textwidth]{figures/screenshot.pdf}
  \end{center} 
  \caption{Screenshot of the ExpressionView Flash applet.} 
  \label{fig:screenshot} 
\end{figure}
\begin{description}
\item[a] Opens an ExpressionView data file. Note that before opening a
  new data file, you should restart the applet, i.e., refresh your
  browser window.
\item[b] Exports the current view to a pdf file. The file also
  includes the title (o) of the gene expression data.
\item[c] Exports the data of the currently viewed module (=bicluster)
  to a CSV file, that can be opened as a spreadsheet.
\item[d] In inspect mode, you can use the mouse to explore the gene
  expression data. The information about the data under the mouse
  pointer is shown in the Info Panel (t).
\item[e, f] Zoom and pan modes allow you to restrict the view to a
  particular part of the gene expression data.
\item[e] In zoom mode, you can also use keyboard shortcuts: {\bf a} to
  auto-zoom onto the modules and {\bf e} to see the whole data. In
  addition to the simple zoom-in feature, you can also use the mouse
  to select the rectangular area you want to have a closer look at.
\item[f] Pan mode.
\item[g, h, i, j] Module highlighting and viewing. It is in general
  impossible to present mutually overlapping biclusters as single
  rectangles. They are made up of a collection of rectangles. The
  ordering algorithm used in the R package realigns the gene
  expression matrix in a way that maximizes the total area of the
  largest rectangle in every bicluster. The outlines of these parts
  are drawn in a slightly brighter color than the background, making
  them easily recognizable.
\item[g, h] Modules are highlighted as the user moves the mouse over
  the gene expression data. The two check boxes allow you to choose
  between highlighting all the parts of a module (Filling) or
  alternatively only the largest rectangle (Outline). You can also
  turn it off completely. For data sets with many modules, it can be
  helpful to restrict highlighting to Outline.
\item[i, j] Similar to the highlighting, these two check boxes allow
  you to show either all the parts of the modules (Filling) or only
  the largest rectangles (Outline). By shift-clicking one of the
  check-boxes you can switch between showing only the modules or only
  the gene expression data.
\item[k] Sets the visibility of the modules layer. Moving the slider
  to the left fades out the gene expression data, thus focusing on the
  Biclusters, while towards the opposite direction, the gene
  expression data moves to the foreground.
\item[l] Realigns the windows at their initial positions.
\item[m] Puts the program in fullscreen mode. Note that due to
  security reasons, it is impossible to enter text in this mode. On
  Mac OS X, a bug in Flash player prevents you from exporting data in
  fullscreen mode.
\item[n] Opens the ExpressionView website, from where you can download
  sample files and tutorials.
\item[o] Description and dimensions of the data set.
\item[p] Modules navigator. The Global tab is always available and
  shows the complete gene expression data. Additional tabs appear as
  you open individual modules. To close a module, simply move the
  mouse over the tab and click the close button that appears. 
\item[q, r, s] Selected genes, samples and modules. The highlighting
  reflects the selection in the tables (w). The selection is
  maintained when switching tabs (p).
\item[q] Selected genes (=probes). 
\item[r] Selected samples (=conditions).
\item[s] Selected modules (=biclusters).
\item[t] Info panel showing the data associated with the current mouse
  position. The GO and KEGG list contain the five most significant
  categories and pathways associated with the modules under the mouse
  pointer.
\item[u] Lists the selected genes, samples and modules, together with
  the intersecting modules.
\item[u1] Opens intersecting modules.
\item[u2] Clears the selection.
\item[v] Lists the selected GO categories and KEGG pathways
\item[w] List navigator. Note that depending on the view (p), the
  lists only show genes and samples contained in the currently viewed
  module. Modules can also be opened by double-clicking on the
  corresponding row. The Experiment tab contains a brief description
  of the data.
\item[x] Searches the tables for a given expression and restricts the
  view to the matching entries. The search function uses Perl-style
  {\bf regular expressions}. By default, the search functions is
  applied to the whole table. To restrict it to a particular column,
  shift-click the corresponding column header.
\item[z] Select a column header to sort the entries according to that
  column. Shift-click to restrict the search function to that column.
\end{description}

\section{Using ExpressionView with non-gene expression data}
While ExpressionView is designed to work with gene expression data
available in the form of a \Rpackage{Bioconductor}
\Rclass{ExpressionSet}, it can also be used to visualize other
data. Let us for instance use in-silico data generated by the
\Rpackage{isa} package with dimensions 50 $\times$ 500 containing 10
overlapping modules:
<<in-silico>>=
library(ExpressionView)
# generate in-silico data with dimensions m x n 
# containing M overlapping modules
# and add some noise
m <- 50
n <- 500
M <- 10
data <- isa.in.silico(num.rows=m, num.cols=n, 
                      num.fact=M, noise=0.1, overlap.row=5)[[1]]
modules <- isa(data)
@

The \Rfunction{ExportEV} uses the named list provided by the
\variable{description} variable to label the data. First, let us
annotate the rows and columns of the data set
<<data set annotation>>=
rownames(data) <- paste("row", seq_len(nrow(data)))
colnames(data) <- paste("column", seq_len(ncol(data)))
@

Next, we assign the meta data associated with the rows of the data
matrix. In this example we use 5 tags labelled ``row tag'':
<<rowdata annotation>>=
rowdata <- outer(1:nrow(data), 1:5, function(x, y) {
  paste("row description (", x, ", ", y, ")", sep="")
})             
rownames(rowdata) <- rownames(data)
colnames(rowdata) <- paste("row tag", seq_len(ncol(rowdata)))
@
And similarly for the columns, using 10 ``column tags'':
<<coldata annotation>>=
coldata <- outer(1:ncol(data), 1:10, function(x, y) {
  paste("column description (", x, ", ", y, ")", sep="")
})
rownames(coldata) <- colnames(data)
colnames(coldata) <- paste("column tag", seq_len(ncol(coldata)))
@

To finish  the description, we add some general information and merge
it with the above tables to get a single named list:
<<in-silico data annotation>>=
description <- list(
experiment=list(
	title="Title", 
	xaxislabel="x-Axis Label",
	yaxislabel="y-Axis Label",
	name="Author", 
	lab="Address", 
	abstract="Abstract", 
	url="URL", 
	annotation="Annotation", 
	organism="Organism"),
coldata=coldata,
rowdata=rowdata
)
@
When dealing with gene expression data, the \variable{xaxislabel} is
equal to ``genes'' and the  \variable{yaxislabel} is
``samples''. Finally, we export the data set to an ExpressionView
file: 
<<in-silico data export, eval=FALSE>>=
ExportEV(modules, data, filename="file.evf", 
         description=description)
@

Simply load this file with the Flash applet and check where the
various labels appear.

\section{Session information}

The version number of R and packages loaded for generating this
vignette were:

<<sessioninfo,results=tex,echo=FALSE>>=
toLatex(sessionInfo())
@

\bibliographystyle{unsrt}
\bibliography{ExpressionView}

\end{document}