### 2010-05-26 expressionviewnote/answers-to-reviewers.tex
EV note: added answers to reviewers.

\documentclass[a4paper]{article}
\usepackage{url}
\usepackage{natbib}
\usepackage{ragged2e}
\usepackage{color}

\newcommand{\Rpackage}[1]{`\texttt{#1}'}
\newcommand{\Rfunction}[1]{`\texttt{#1}'}
\newcommand{\Rclass}[1]{\textsl{#1}}

\newenvironment{myquote}{\begin{quote}\color{blue}}{\end{quote}}

\setlength{\parskip}{12pt}
\setlength{\parindent}{0pt}

\title{ExpressionView -- an interactive viewer for
  modules identified in gene expression data\\[10pt]
  Answers to reviewers}
\author{Andreas L\"uscher, G\'abor Cs\'ardi, \\ Aitana Morton de
  Lachapelle, Zolt\'an Kutalik, \\ and Sven Bergmann}

\begin{document}

\maketitle

\RaggedRight

First of all, we want to thank the reviewers for their thorough work
and useful suggestions. In the following, the reviewers' comments are
typeset in blue, and indented as a block.

\section*{Reviewer 1}

\begin{myquote}
All comments are for minor revisions:

1) BiVisu by Cheng et al shall be cited:
Bioinformatics. 2007;23(17):2342-4. Please briefly elaborate the major
difference(s) of ExpressionView from BiVisu and BicOverlapper.
\end{myquote}

We have added a brief comparison to Bivisu~\citep{cheng07} and
BicOverlapper\citep{santamaria08}.

\begin{myquote}
2) Page 2, line 30: please re-write the sentence "We find that in
almost every situation, the proposed algorithm finds an order that
recovers more than 99\% of the score of the optimal solution and in
most cases, it recovers the correct alignment". 
\end{myquote}

We have rewritten this part.

\begin{myquote}
3) Page 2, line 55, change "any computer" to "a computer".
\end{myquote}

We have corrected this.

\section*{Reviewer 2}

\begin{myquote}
Comments to the Author

This is a very useful tool which users might  benefit if there is
clearly stated in the website, what sort of input format the
application accepts. There does not seem to be a complete and explicit
description of the various types of input format beyond just a
expression matrix.
\end{myquote}

For the matrix reordering, ExpressionView accepts modules in the
format defined in the eisa R package~\citep{csardi10}
(\textit{ISAModules} object) and the
Biclust R package~\citep{kaiser08} (\textit{Biclust} object). These
cover the existing R implementations of biclustering algorithms (to our
knowledge). Moreover, the Biclust package defines a very simple class
for storing biclustering results, so the output of any other algorithm
(implemented in R or not) can be easily transformed into this
format. This is documented in the Biclust package.

For the expression data itself, ExpressionView accepts
\textit{ExpressionSet} objects as the input. This is the standard data
type in BioConductor for storing expression data. E.g. the GCRMA and
other normalization methods all output this format and it is also easy
to create an \textit{ExpressionSet} object from a simple tabular text
file.

Alternatively, ExpressionView can visualize non-expression data,
inputted as a simple matrix, this is explained in the R manual that
comes with the package.

\begin{myquote}
For example, how do we incorporate GO terms and
KEGG pathways into the raw input dataset, or are these also computed
by the application? 
\end{myquote}

These are computed by the R part of the application, using functions
in the eisa R package. There are examples for this in the manual that
comes with the package.

\begin{myquote}
It is not entirely clear to the casual reader,
whether the Flash application takes raw data or the intermediate XML
result from the workflow, which means user may need to use command
line R anyway to generate the data. If it does so, the format of this
needs to be clearly stated for a non-bioinformatics biologist to be
able to load and run the application. 
\end{myquote}

The Flash application takes the intermediate XML file, we made this
clear in Fig.~1. The idea is that the user performing the analysis in
R is usually not the same, as the user browsing the results. So the
(bioinformatician) user generates the data in R, exports it into an
XML file, and the other (life scientist) user explores it using an
ordinary web browser. We have made this distinction clear in the
manuscript.

The file format of the XML file is not relevant in most cases, as the
file is produced by the ExpressionView package, and the user does not
have to produce it by hand. It is only relevant, if the user wants to
import data that was produced by another reordering algorithm, into
the Flash applet. We have created an XML Schema document that precisely
defines the XML data file's format, and also written a document that
discusses the format. These are available on the ExpressionView website.

\begin{myquote}
The tutorials are impressive, but the audio volume could be increased
or spoken by a person with a higher pitch voice. Very difficult to
hear.  
\end{myquote}

We have increased the audio volume in the video tutorials.

\begin{myquote}
ExpressionView source code (for the Flash applet) in winrar
format contains a lot of empty folders and tmp files some of which may
not be necessary. 
\end{myquote}

We have simplified the directory structure of the source code.

\begin{myquote}
The paper could do with at least a few words in the body of the paper
or in the website, to explain what improvement over features of other
existing Expression data biclustering visualisers currently
available. 
\end{myquote}

We have included a brief comparison to Bivisu~\citep{cheng07} and
BiocOverlapper~\citep{santamaria08}, two existing bicluster
visualization software packages.

\begin{myquote}
Moreover, the paper must a least emphasize justification
for why biologists should find this application/algorithm more useful
over other existing applications/algorithms besides just visual
appeal. 
\end{myquote}

We emphasized the advantages of ExpressionView more in the paper:
\begin{itemize}
\item visualization of all the modules and the expression values together.
\item ease of use for the (life scientist) end user, only a web browser is
  needed, no extra programs needed to be installed.
\item the inclusion of meta-data and enrichment data into the same
  environment, with interactive links to corresponding databases.
\end{itemize}

\section*{Reviewer 3}

\begin{myquote}
Comments to the Author

General comments:

Luescher et al. describe software for visualizing modules in gene
expression data using an efficient implementation of matrix
reordering. Such a package is of high practical value for biologists
to analyze and visualize gene expression data, and I therefore
recommend publication as an Applications Note. In particular, I
appreciate the connection to the Bioconductor package, it is well
done.

Specific comments:

Minor:

I would recommend a different name for the package, ExpressionView is
too general.
\end{myquote}

While this is just criticism, we cannot easily change the name of the
package, as it is part of BioConductor now, and the name change would
cause a lot problems for the users and the BioConductor developers.

\begin{myquote}
Biologists should have no hurdles to access and apply the software
\ldots I could not easily install it on my LINUX/Ubuntu computer after
having installed the isa2 package, as recommended by the authors. 
I assume the download and installation of the package will be as
straightforward as the isa2 package by the authors.
\end{myquote}

The ExpressionView package is an official BioConductor package since
April 2010, so its installation should be easy on any platforms
supported by BioConductor (Linux, MS Windows, Apple OSX). We have
updated the installation instructions on the ExpressionView homepage
accordingly.

\begin{myquote}
The authors should consider that some computers might not play the
Flash applets, and should provide short and clear instructions for
installation on their website.
\end{myquote}

We have added a note about this, and a link to the relevant Adobe
website, to the ExpressionView homepage.

\bibliographystyle{natbib}
\bibliography{expressionview}

\end{document}